---
title: "RAIS scripts"
author: "Tainá Souza Pacheco"
date: ''
output: 
  html_document:
    theme: united
    toc: true
    number_sections: true
    toc_depth: 4 
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: hide
---

<style>
div.orange { background-color:#fee6ce; border-radius: 5px; padding: 20px;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

This document comments the scripts that generated the data used in the paper.

# RAIS structure

RAIS (*relação anual de informações sociais*) is an annual register of the formal labor market in Brazil, both private and public sectors. All employees that had at least one formal job in the year must appear.

The **data is structure at individual-occupation level**, so one person can appear more than once each year. This can happens if:

-   The person had two simultaneous job
    -   for instance a teacher that teach in two different schools, one in the morning and other in the afternoon, or
    -   a public servant that accumulates functions.
-   The person had have two non-simultaneous jobs in that year:
    -   the person had one occupation in one company, quit/was fired and then found another occupation in the same year.

If the person quit/was fired in one year but only found a new job in a subsequent year, then we only see the person once.

# Unique identifier

RAIS has two unique individual identifier: CPF (*cadastro único de pessoa física*) and PIS (*programa de integração social*). The first was only included in RAIS in 2002. Since we are looking into data from 1998, we had to relly on the PIS number to uniquely identify individuals.

For companies the unique identifier is CNPJ (*cadastro nacional de pessoa jurídica*).

# **2_matching_public_private_v6.R**

```{r full code matching, eval = FALSE}
######################
#### CBo matching ####
######################
# 1) Para os anos de 2001, 2002, 2003 e 2004 (2 antes e 2 depois de Lula) classificar todas as observações em público
# ou privado com base em TODAS as Naturezas Jurídicas;
# 2) Para a base de dados classificada, filtrar para CBO privado OU público E federal;
# 3) Gerar uma lista com CBO públicos federais para esses anos;
# 4) Manualmente classificar os CBOs públicos federais que NÃO PODEM ser encontrados no setor público (ex. juiz);
# 5) Excluir da base CBO que são públics e foram classificados como privado. 
# Decidimos excluir porque não sabemos se o erro foi no preenchimento do CBO ou da Natureza Jurídica;
# 6) Com essa base limpa, fazer o match ano a ano. 
# Esse match vai buscar CBOs que em um mesmo ano foram encontrados tanto no setor público quando no setor privado.
# Exemplos: professores, enfermeiros, advogados, economistas...;
# 7) Fazer uma lista dos CBOs classificados como público que não tiveram match;
# 8) Para os CBOs de 7), buscar outras posições das pessoas nos três anos anteriores (menor ano será 1998) no setor privado.


# colunas do banco de dados a nível individual
    # PIS >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>> ()
    # CBO >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>> ()
    # NatJur >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>> ()
    # CNPJ raiz/radical >>>>>>>>>>>>>>>>>>>>> ()
    # salario contratual >>>>>>>>>>>>>>>>>>>> ()
    # remuneração media >>>>>>>>>>>>>>>>>>>>> ()
    # remuneração em dez >>>>>>>>>>>>>>>>>>>> ()
    # última remuneração >>>>>>>>>>>>>>>>>>>> ()
    # horas de trabalho contratual >>>>>>>>>> ()
    # data início do contrato >>>>>>>>>>>>>>> ()
    # tipo de admissão >>>>>>>>>>>>>>>>>>>>>> ()
    # data/mês fim do contrato >>>>>>>>>>>>>> ()
    # status do emrpego em 31/dez >>>>>>>>>>> ()
    # causa do desligamento >>>>>>>>>>>>>>>>> ()
    # tipo de contrato >>>>>>>>>>>>>>>>>>>>>> ()
    # data de nascimento >>>>>>>>>>>>>>>>>>>> ()
    # gênero >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>> ()
    # grau de instrução >>>>>>>>>>>>>>>>>>>>> ()
    # tempo de emprego >>>>>>>>>>>>>>>>>>>>>> ()

rm(list=ls())
gc()

library("tidyverse")
library("readxl")
library("data.table")

# Para facilitar
select <- dplyr::select
filter <- dplyr::filter
`%notin%` = Negate(`%in%`)

# personal_directory <- "//fs-eesp-01/EESP/Usuarios/taina.pacheco/"
# directory <- "//fs-eesp-01/eesp-bd-02/RAIS_IDENTIFICADA/"

personal_directory <- "C:/Users/taina/Desktop/RAIS/matching/dec2021/"
directory <- "C:/Users/taina/Desktop/RAIS/"

#years
years_lula <- as.character(c(2001:2004))
years <- as.character(c(1995:2004))

#http://www.mtecbo.gov.br/
nat_jur_fed <- as.character(c(1015, 1040, 1074, 1104, 1139, 1163, 1252, 10))
nat_jur <- as.character(c(10,
                          # Natureza Jurídica 1995
                          # https://concla.ibge.gov.br/estrutura/natjur-estrutura/natureza-juridica-1995.html
                          1015, #Poder Executivo Federal
                          1023, #Poder Executivo Estadual
                          1031, #Poder Executivo Municipal
                          1040, #Poder Legislativo Federal
                          1058, #Poder Legislativo Estadual
                          1066, #Poder Legislativo Municipal
                          1074, #Poder Judiciario Federal
                          1082, #Poder Judiciario Estadual
                          1090, #Orgão Autonomo de Direito Público
                          1104, #Autarquia Federal
                          1112, #Autarquia Estadual
                          1120, #Autarquia Municipal
                          1139, #Fundação Federal
                          1147, #Fundação Estadual
                          1155, #Fundação Municipal
                          1996, #Outras Formas de Organização da Administração Pública
                          2011, #Empresa Publica - Sociedade por Quotas de Responsabilidade Limitada
                          2020, #Empresa Publica - Sociedade Anonima de Capital Fechado
                          2038, #Sociedade Anonima de Capital Aberto com Controle Acionário Estatal (sociedade de economia mista) 
                          
                          # Natureza jurídica 2002
                          # https://concla.ibge.gov.br/estrutura/natjur-estrutura/natureza-juridica-2002.html
                          1015, #Órgão Público do Poder Executivo Federal
                          1023, #Órgão Público do Poder Executivo Estadual ou do Distrito Federal
                          1031, #Órgão Público do Poder Executivo Municipal
                          1040, #Órgão Público do Poder Legislativo Federal
                          1058, #Órgão Público do Poder Legislativo Estadual ou do Distrito Federal
                          1066, #Órgão Público do Poder Legislativo Municipal
                          1074, #Órgão Público do Poder Judiciário Federal
                          1082, #Órgão Público do Poder Judiciário Estadual
                          1104, #Autarquia Federal
                          1112, #Autarquia Estadual ou do Distrito Federal
                          1120, #Autarquia Municipal
                          1139, #Fundação Federal
                          1147, #Fundação Estadual ou do Distrito Federal
                          1155, #Fundação Municipal
                          1163, #Órgão Público Autônomo da União
                          1171, #Órgão Público Autônomo Estadual ou do Distrito Federal
                          1180, #Órgão Público Autônomo Municipal
                          2011, #Empresa Pública
                          2038)) #Sociedade de Economia Mista 

# There were changes in Natureza Jurídica after 2002, but none of them altering the classification of public entities.
# https://concla.ibge.gov.br/classificacoes/por-tema/organizacao-juridica
                        
open_files <- function(years_files = NA_character_, file_to_read = NA_character_){
  if(years_files == "2002"){
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("OCUPACAO", "PIS", "NATUR JUR",
                           "HORAS CONTR", "REM DEZ (R$)", "REM MED (R$)", "ULT REM", "SAL CONTR"
                           # ,"DT ADMISSAO", "RADIC CNPJ", "GRAU INSTR", "MES DESLIG", "CAUSA DESLI", "MUNICIPIO",
                           # "SEXO", "TEMP EMPR", "TP VINCL", "EMP EM 31/12", "DT NASCIMENT"
                           )) %>% 
      rename(natjur = `NATUR JUR`, 
             cbo = `OCUPACAO`,
             # date_adm = `DT ADMISSAO`,
             # CNPJ_root = `RADIC CNPJ`,
             # educ = `GRAU INSTR`,
             # month_quit = `MES DESLIG`,
             # reason_quit = `CAUSA DESLI`,
             # municipality = MUNICIPIO,
             # gender = SEXO,
             # time_employed = `TEMP EMPR`,
             # tp_contract = `TP VINCL`,
             # employed_dec = `EMP EM 31/12`,
             # b_date = `DT NASCIMENT`,
             hours_contract = `HORAS CONTR`,
             wage_dec = `REM DEZ (R$)`,
             wage_mean = `REM MED (R$)`,
             wage_last = `ULT REM`,
             wage_contract = `SAL CONTR`) %>% 
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             # age = as.numeric(years_files) - as.numeric(substr(b_date,(nchar(b_date)+1)-4,nchar(b_date))),
             # date_adm = ifelse(nchar(date_adm) == 7, paste0("0", date_adm), date_adm),
             # date_adm = as.Date(date_adm, format = "%d%m%Y"),
             # date_quit = ifelse(month_quit == "00", NA, 
             #                    as.Date(paste0("30/", month_quit, "/", years_files), format = "%d/%m/%Y")),
             # wage_dec = as.numeric(str_replace(wage_dec, ",", ".")),
             wage_mean = as.numeric(str_replace(wage_mean, ",", ".")),
             wage_last = as.numeric(str_replace(wage_last, ",", ".")),
             wage_contract = as.numeric(str_replace(wage_contract, ",", ".")),
             wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                                as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) #%>% 
      # select(-b_date, -month_quit)
      # filter(cbo %in% c("20496", "20512", "20528", "20544",
      #                     "21420", "21490", 
      #                     "24592", "24608", "24624", "24640", 
      #                     "16416", "16432", "16448", 
      #                     "28688", "28704", 
      #                     "39385"))

  } else if (years_files %in% as.character(1995:1998)){
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("CBO 94 Ocupação", "PIS", "Natureza Jurídica", 
                           "Vl Remun Dezembro (SM)", "Vl Remun Média (SM)", "Qtd Hora Contr"
                           # ,"Ano Admissão", "CNPJ Raiz", "Grau Instrução 2005-1985", "Mês Admissão",
                           # "Mês Desligamento", "Motivo Desligamento", "Município", "Sexo Trabalhador", "Tempo Emprego",
                           # "Tipo Vínculo", "Vínculo Ativo 31/12", "Idade", "Tipo Admissão"
                )) %>% 
      rename(natjur = `Natureza Jurídica`, 
             cbo = `CBO 94 Ocupação`,
             # year_adm = `Ano Admissão`,
             # CNPJ_root = `CNPJ Raiz`,
             # educ = `Grau Instrução 2005-1985`,
             # month_adm = `Mês Admissão`,
             # month_quit = `Mês Desligamento`,
             # reason_quit = `Motivo Desligamento`,
             # municipality = `Município`,
             # gender = `Sexo Trabalhador`,
             # time_employed = `Tempo Emprego`,
             # tp_contract = `Tipo Vínculo`,
             # employed_dec = `Vínculo Ativo 31/12`,
             # age = `Idade`,
             hours_contract = `Qtd Hora Contr`,
             wage_dec = `Vl Remun Dezembro (SM)`,
             wage_mean = `Vl Remun Média (SM)`) %>% 
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             wage_last = NA_character_,
             wage_contract = NA_character_,
             wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                                as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) # %>% 
    # select(-month_adm, -year_adm, -month_quit)
    # filter(cbo %in% c("20496", "20512", "20528", "20544",
    #                     "21420", "21490", 
    #                     "24592", "24608", "24624", "24640", 
    #                     "16416", "16432", "16448", 
    #                     "28688", "28704", 
    #                     "39385"))
    
  } else if (years_files %in% as.character(1995)){
    df <- df %>% 
      mutate(# Maio 1995,  Lei nº 9032, de 1995, salário mínimo: R$ 100,00
        wage_dec = as.numeric(str_replace(wage_dec, ",", ".")) * 100,
        wage_mean = as.numeric(str_replace(wage_mean, ",", ".")) * 100,
        wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                           as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) 
  } else if (years_files %in% as.character(1996)){
    df <- df %>% 
      mutate(# Maio 1996,  Lei nº 9971, de 1996, salário mínimo: R$ 120,00
        wage_dec = as.numeric(str_replace(wage_dec, ",", ".")) * 112,
        wage_mean = as.numeric(str_replace(wage_mean, ",", ".")) * 112,
        wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                           as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) 
  } else if (years_files %in% as.character(1997)){
    df <- df %>% 
      mutate(# Maio 1997,  Lei nº 9971, de 2000, salário mínimo: R$ 130,00
        wage_dec = as.numeric(str_replace(wage_dec, ",", ".")) * 120,
        wage_mean = as.numeric(str_replace(wage_mean, ",", ".")) * 120,
        wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                           as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) 
  } else if (years_files %in% as.character(1998)){
    df <- df %>% 
      mutate(# Maio 1998,  Lei nº 9971, de 2000, salário mínimo: R$ 130,00
             wage_dec = as.numeric(str_replace(wage_dec, ",", ".")) * 130,
             wage_mean = as.numeric(str_replace(wage_mean, ",", ".")) * 130,
             wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                                as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) 
  } else if (years_files %in% as.character(c(1999:2001))){
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("CBO 94 Ocupação", "PIS", "Natureza Jurídica", 
                           "Vl Remun Dezembro Nom", "Vl Remun Média Nom", "Qtd Hora Contr"
                           # ,"Ano Admissão", "CNPJ Raiz", "Grau Instrução 2005-1985", "Mês Admissão",
                           # "Mês Desligamento", "Motivo Desligamento", "Município", "Sexo Trabalhador", "Tempo Emprego",
                           # "Tipo Vínculo", "Vínculo Ativo 31/12", "Idade", "Tipo Admissão"
                           )) %>% 
      rename(natjur = `Natureza Jurídica`, 
             cbo = `CBO 94 Ocupação`,
             # year_adm = `Ano Admissão`,
             # CNPJ_root = `CNPJ Raiz`,
             # educ = `Grau Instrução 2005-1985`,
             # month_adm = `Mês Admissão`,
             # month_quit = `Mês Desligamento`,
             # reason_quit = `Motivo Desligamento`,
             # municipality = `Município`,
             # gender = `Sexo Trabalhador`,
             # time_employed = `Tempo Emprego`,
             # tp_contract = `Tipo Vínculo`,
             # employed_dec = `Vínculo Ativo 31/12`,
             # age = `Idade`,
             hours_contract = `Qtd Hora Contr`,
             wage_dec = `Vl Remun Dezembro Nom`,
             wage_mean = `Vl Remun Média Nom`)%>% 
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             # date_adm = as.Date(paste0("01/", month_adm, "/", year_adm), format = "%d/%m/%Y"),
             # date_quit = ifelse(month_quit == "00", NA, 
             #                    as.Date(paste0("30/", month_quit, "/", years_files), format = "%d/%m/%Y")),
             wage_dec = as.numeric(str_replace(wage_dec, ",", ".")),
             wage_mean = as.numeric(str_replace(wage_mean, ",", ".")),
             wage_last = NA_character_,
             wage_contract = NA_character_,
             wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                                as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) #%>% 
      # select(-month_adm, -year_adm, -month_quit)
      # filter(cbo %in% c("20496", "20512", "20528", "20544",
      #                     "21420", "21490", 
      #                     "24592", "24608", "24624", "24640", 
      #                     "16416", "16432", "16448", 
      #                     "28688", "28704", 
      #                     "39385"))
    
  } else{
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("OCUPACAO 94", "PIS", "NATUR JUR", 
                           "HORAS CONTR", "REM DEZ (R$)", "REM MED (R$)", "ULT REM", "SAL CONTR"
                           # ,"DT ADMISSAO", "RADIC CNPJ", "GRAU INSTR", "MES DESLIG", "CAUSA DESLI", 
                           # "MUNICIPIO", "SEXO", "TEMP EMPR", "TP VINCULO", "EMP EM 31/12", "DT NASCIMENT"
                           )) %>% 
      rename(natjur = `NATUR JUR`, 
             cbo = `OCUPACAO 94`,
             # date_adm = `DT ADMISSAO`,
             # CNPJ_root = `RADIC CNPJ`,
             # educ = `GRAU INSTR`,
             # month_quit = `MES DESLIG`,
             # reason_quit = `CAUSA DESLI`,
             # municipality = MUNICIPIO,
             # gender = SEXO,
             # time_employed = `TEMP EMPR`,
             # tp_contract = `TP VINCULO`,
             # employed_dec = `EMP EM 31/12`,
             # b_date = `DT NASCIMENT`,
             hours_contract = `HORAS CONTR`,
             wage_dec = `REM DEZ (R$)`,
             wage_mean = `REM MED (R$)`,
             wage_last = `ULT REM`,
             wage_contract = `SAL CONTR`) %>% 
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             # age = as.numeric(years_files) - as.numeric(substr(b_date,(nchar(b_date)+1)-4,nchar(b_date))),
             # date_adm = ifelse(nchar(date_adm) == 7, paste0("0", date_adm), date_adm),
             # date_adm = as.Date(date_adm, format = "%d%m%Y"),
             # date_quit = ifelse(month_quit == "00", NA, 
             #                    as.Date(paste0("30/", month_quit, "/", years_files), format = "%d/%m/%Y")),
             wage_dec = as.numeric(str_replace(wage_dec, ",", ".")),
             wage_mean = as.numeric(str_replace(wage_mean, ",", ".")),
             wage_last = as.numeric(str_replace(wage_last, ",", ".")),
             wage_contract = as.numeric(str_replace(wage_contract, ",", ".")),
             wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                                as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) #%>% 
      # select(-b_date, -month_quit)
    # filter(cbo %in% c("20496", "20512", "20528", "20544",
    #                       "21420", "21490", 
    #                       "24592", "24608", "24624", "24640", 
    #                       "16416", "16432", "16448", 
    #                       "28688", "28704", 
    #                       "39385"))
  }
  
  return(df)
}

list_cbo <- list()
for(i in seq_along(years)){
  
  print(years[i])
  
  # list files for each year
  files <- grep(list.files(path = paste0(directory, years[i]), 
                           all.files = FALSE,
                           full.names = TRUE), 
                pattern= ('Estb|ESTB|IGNORADO'), 
                inv=TRUE, value=TRUE)
  
  list_data_temp <- list()
  list_cbo_temp <- list()
  for(j in seq_along(files)) { 
    # 1) Para os anos de 2001, 2002, 2003 e 2004 (2 antes e 2 depois de Lula) classificar todas as 
    # observações em público ou privado com base em TODAS as Naturezas Jurídicas;
    list_data_temp[[j]] <- open_files(years_file = years[i], file_to_read = files[j]) %>% 
      mutate(public = 0, public = ifelse(natjur %in% nat_jur, 1, public),
             federal = 0, federal = ifelse(natjur %in% nat_jur_fed, 1, federal)) %>% 
    # 2) Para a base de dados classificada, filtrar para CBO privado OU público E federal;
      filter(public == 0 | federal == 1) %>% 
    # criar marcação de ano
      mutate(year = years[i])
    
    # 3) Gerar uma lista com CBO públicos federais para esses anos;
    list_cbo_temp[[j]] <- list_data_temp[[j]] %>% 
      filter(federal == 1) %>% 
      distinct(cbo) %>% 
      select(cbo)
    
    print(j)
  }
  
  df <- bind_rows(list_data_temp)
  saveRDS(df, paste0(personal_directory, "list_data_wages/data_wages_", years[i], ".RDS"))
  
  list_cbo[[i]] <- bind_rows(list_cbo_temp) %>% distinct(cbo)
  
  remove(df, list_data_temp, list_cbo_temp)
}

# 4) Manualmente classificar os CBOs públicos federais que NÃO PODEM ser encontrados no setor público (ex. juiz);
cbo_federais <- bind_rows(list_cbo) %>% distinct(cbo)

# write.csv(cbo_federais, file = paste0(personal_directory, "cbos_federal.csv"))

remove(list_cbo)
# CBOs que são apenas públicos:
# No CBO 94 não havia presidente, por exemplo. Está dentro de dirigentes do setor público
# 31920 >	Agente de saúde pública
# 21420 >	Funcionário público federal superior (1114-05 no CBO 2002)
# 31490 >	Outros serventuários da justiça e trabalhadores assemelhados
# 31390 >	Outros agentes superiores de polícia
# 21490 >	Outros funcionários públicos superiores
# 21320 >	Juiz federal
# 31440 >	Oficial de justiça
# 21330 >	Juiz estadual
# 31430 >	Escrivão
# 58290 >	Outros policiais e trabalhadores assemelhados
# 31420 >	Tabelião
# 21390 >	Outros membros superiores do Poder Judiciário
# 21350 >	Juiz do trabalho
# 21340 >	Juiz militar
# 21220 >	Membro superior do Poder Executivo
# 21130 >	Deputado federal
# 21360 >	Juiz eleitoral
# 31320 >	Delegado de polícia
# 21120 >	Senador
# 21140 >	Vereador
# 21430 >	Funcionário público estadual superior
# 12920 >	Procurador da fazenda nacional
# 58250 >	Papiloscopista policial
# 31340 >	Perito criminal
# 58430 >	Policial rodoviário
# 39385 >	Despachante aduaneiro
# 58230 >	Detetive de polícia
cbo_only_public <- as.character(c(31920, 21420, 31490, 31390, 21490, 21320, 31440, 21330, 31430, 
                                  58290, 31420, 21390, 21350, 21340, 21220, 21130, 21360, 31320, 
                                  21120, 21140, 21430, 12920, 58250, 31340, 58430, 39385, 58230))

# 5) Excluir da base CBO que são públics e foram classificados como privado. 
# Decidimos excluir porque não sabemos se o erro foi no preenchimento do CBO ou da Natureza Jurídica;
# list_data <- readRDS(paste0(personal_directory, "list_data.RDS"))

files <- list.files(path = paste0(personal_directory, "list_data_wages/"), 
                    all.files = FALSE,
                    full.names = TRUE)

for (i in seq_along(files)) {
  df <- readRDS(files[i])
  
  print(years[i])
  print(nrow(df))
  # os CBOs que não estão na lista podem ser públicos e privados
  df1 <- df %>% filter(cbo %notin% cbo_only_public)
  
  # para os CBOs que estão na lista só vamos manter as observações corretamente classificadas como públicas
  df2 <- df %>% 
    filter(cbo %in% cbo_only_public) %>% 
    filter(federal == 1)
  
  df <- rbind(df1, df2)
  print(nrow(df))
  
  saveRDS(df, paste0(personal_directory, "/list_data_wages_clean/data_wages_clean_", years[i], ".RDS"))
  
  remove(df, df1, df2)
}

# Número de observações que perdemos por ano:
# 1995: 27,562,711 - 27,541,093 = 21,618
# 1996: 26,516,841 - 26,493,344 = 23,497
# 1997: 27,863,657 - 27,837,125 = 26,532
# 1998: 28,194,129 - 28,158,249 = 35,880
# 1999: 28,460,449 - 28,404,008 = 56,441
# 2000: 30,473,813 - 30,428,651 = 45,162
# 2001: 32,019,990 - 31,966,857 = 53,133
# 2002: 36,986,619 - 36,914,780 = 71,839
# 2003: 33,930,145 - 33,878,305 = 51,840
# 2004: 36,376,481 - 36,325,087 = 51,394

# list_data <- readRDS(paste0(personal_directory, "list_data_wages_clean.RDS"))
# 6) Com essa base limpa, fazer o match ano a ano. 
files <- list.files(path = paste0(personal_directory, "list_data_wages_clean/"), 
                    all.files = FALSE,
                    full.names = TRUE)

list_cbo_matched <- list()
list_cbo_not_matched <- list()
for (i in seq_along(files)) {
  print(years[i])
  
  df <- readRDS(files[i])
  
  cbos_duplicates <- df %>% 
    count(cbo, public) %>% 
    add_count(cbo, name = "duplicates") %>% 
    filter(duplicates > 1) %>% 
    ungroup() %>% 
    distinct(cbo) %>% 
    pull(cbo)
  
  df_cbo_matched <- df %>% 
    filter(cbo %in% cbos_duplicates) %>% 
    group_by(cbo, public) %>% 
    mutate(observations = n(),
           wage_mean_mean = mean(as.numeric(wage_mean), na.RM = TRUE),
           wage_mean_median = median(as.numeric(wage_mean), na.RM = TRUE),
           # wage_mean_var = ifelse(n %in% c(1, 2), 0, var(as.numeric(wage_mean), na.RM = TRUE)),
           wage_mean_min = min(as.numeric(wage_mean), na.RM = TRUE),
           wage_mean_max = max(as.numeric(wage_mean), na.RM = TRUE),
           hours_contract = mean(as.numeric(hours_contract), na.rm = TRUE),
           wage_hour_mean = mean(as.numeric(wage_hour), na.rm = TRUE),
           year = years[i]) %>% 
    ungroup() 
  
  saveRDS(df_cbo_matched, 
          paste0(personal_directory, "list_cbo_direct_match_wages/cbo_direct_match_wages_", years[i], ".RDS"))
  
  df_cbo_no_matched <- df %>% 
    filter(cbo %notin% cbos_duplicates) %>% 
    filter(public == 1) %>% 
    mutate(year = years[i]) %>% 
    ungroup()
  
  saveRDS(df_cbo_no_matched, 
          paste0(personal_directory, "list_cbo_no_direct_match_wages/cbo_no_direct_match_wages_", years[i], ".RDS"))
  
  remove(df, df_cbo_matched, df_cbo_no_matched)
    
}

files <- list.files(path = paste0(personal_directory, "list_cbo_direct_match_wages/"), 
                    all.files = FALSE,
                    full.names = TRUE)

for(i in seq_along(files)){
  df <- readRDS(files[i])
  
  df <- df %>% 
    filter(federal == 1)
  
  saveRDS(df, 
          paste0(personal_directory, "list_cbo_direct_match_wages_federal/cbo_direct_match_wages_federal_", 
                 years[i], ".RDS"))
}

# 7) Fazer uma lista dos CBOs classificados como público que não tiveram match;
files <- list.files(path = paste0(personal_directory, "list_cbo_no_direct_match_wages/"), 
                    all.files = FALSE,
                    full.names = TRUE)

vec_cbo_no_direct_match <- c()
for(i in seq_along(files)){
  df <- readRDS(files[i])
  
  df <- df %>% 
    filter(federal == 1)
  
  vec_cbo_no_direct_match <- c(vec_cbo_no_direct_match, 
                               (df %>% distinct(cbo) %>% pull(cbo)))
  
  saveRDS(df, 
          paste0(personal_directory, "list_cbo_no_direct_match_wages_federal/cbo_no_direct_match_wages_federal_", 
                 years[i], ".RDS"))
}

vec_cbo_no_direct_match <- unique(vec_cbo_no_direct_match)

# 8) Para os CBOs de 7), buscar outras posições das pessoas nos três anos anteriores (menor ano será 1995) no setor privado.
files <- list.files(path = paste0(personal_directory, "list_cbo_no_direct_match_wages_federal/"), 
                    all.files = FALSE,
                    full.names = TRUE)

files_bsearch <- list.files(path = paste0(personal_directory, "list_data_wages_clean/"), 
                    all.files = FALSE,
                    full.names = TRUE)

rename_fun <- function(x){
  x <- paste0(x, "2")
  
  return(x)
}

sub_comma_fun <- function(str){
  str <- gsub(",", ".", str)
  
  return(str)
}

inflation_index <- read_csv(paste0(personal_directory, "input/INPC_correction.csv"), 
                            col_types = cols(INPC = col_number(), 
                                             year_destination = col_character(), 
                                             year_origin = col_character()))
for(i in length(files):4){
  df <- readRDS(files[i])
  
  CBO_vector <- df %>% 
    distinct(cbo) %>% 
    pull(cbo)
  
  df <- df %>% 
    mutate_at(c("hours_contract", "wage_dec", "wage_mean", "wage_last", "wage_contract", 
                "public", "federal", "wage_hour"), as.numeric) %>% 
    mutate_at(c("cbo", "PIS", "natjur", "year"), as.character) %>% 
    filter(cbo %in% CBO_vector) %>% 
  # excluir PIS "errados"
    filter(PIS %notin% c("00000000000", "0", "99999999999"))
  
  PIS_vector <- df %>% 
    filter(PIS != "0") %>% 
    pull(PIS)

  list_aux <- list()
  PIS_matched <- c()
  j <- 1
  for (k in c(i - 1, i - 2, i - 3)) { #starts at year - 1
    if(k <= 0){
      print(k)
    } 
    else {
      df_bsearch <- readRDS(files_bsearch[k]) %>% 
        # excluir PIS "errados"
        filter(PIS %notin% c("00000000000", "0", "99999999999"))
      
      inflation <- inflation_index %>% 
        filter(year_destination == years[i],
               year_origin == years[k]) %>% 
        pull(INPC)
      
      list_aux[[j]] <- df_bsearch %>% 
        filter(public == 0) %>% 
        filter(PIS %in% PIS_vector) %>% 
        mutate_at(c("wage_dec", "wage_mean", "wage_last", "wage_contract"), sub_comma_fun) %>% 
        mutate_at(c("hours_contract", "wage_dec", "wage_mean", "wage_last", "wage_contract", 
                    "public", "federal", "wage_hour"), as.numeric) %>% 
        mutate_at(c("cbo", "PIS", "natjur", "year"), as.character) %>% 
        mutate(wage_dec = wage_dec*inflation, 
               wage_mean = wage_mean*inflation, 
               wage_last = wage_last*inflation, 
               wage_contract = wage_contract*inflation,
               wage_hour = wage_hour*inflation)
        
      PIS_matched <- c(PIS_matched, 
                       (list_aux[[j]] %>% pull(PIS)))
      
      PIS_vector <- PIS_vector[PIS_vector %notin% PIS_matched]
      
      j <- j + 1
      
      remove(df_bsearch)
    }
  }
  df2 <- bind_rows(list_aux) %>% 
    distinct(PIS, cbo, hours_contract, wage_contract, wage_hour, year, .keep_all = TRUE)
  
  df <- df %>% 
    select(PIS, cbo) %>% 
    rename(cbo_original = cbo) %>%
    left_join(df2, by = "PIS")
  
  df_matched <- df %>% 
    filter(!is.na(cbo)) %>% 
    add_count(cbo_original, cbo)
  
  saveRDS(df_matched, 
          paste0(personal_directory, 
                 "list_cbo_match_wages_public_past_years/cbo_match_wages_public_past_years_", 
                 years[i], ".RDS"))
  
  df_n_matched <- df %>% 
    filter(is.na(cbo))
  
  saveRDS(df_n_matched, 
          paste0(personal_directory, 
                 "list_cbo_n_match_wages_public_past_years/list_cbo_n_match_wages_public_past_years_", 
                 years[i], ".RDS"))
  
  print(i)
  
  remove(df, df_matched, df_n_matched)
}

# 8a) padronizar o banco de dados para juntar com os outros depois
files_matched <- list.files(path = paste0(personal_directory, "list_cbo_match_wages_public_past_years/"), 
                            all.files = FALSE,
                            full.names = TRUE)

files_clean <- list.files(path = paste0(personal_directory, "list_cbo_no_direct_match_wages_federal/"), 
                          all.files = FALSE,
                          full.names = TRUE)

files_clean <- files_clean[4:length(files_clean)]

for (i in seq_along(files_matched)) {
  df_matched <- readRDS(files_matched[i]) %>% 
    select(PIS, cbo_original, wage_mean, hours_contract, wage_hour) %>% 
    rename(wage_mean_mean = wage_mean,
           wage_hour_mean = wage_hour,
           hours_contract_mean = hours_contract,
           cbo = cbo_original) %>% 
    mutate(wage_mean_median = NA_integer_,
           wage_mean_min = NA_integer_,
           wage_mean_max = NA_integer_) 
  
  PIS_matched <- df_matched %>% 
    distinct(PIS) %>% pull(PIS)
  
  df_clean <- readRDS(files_clean[i]) %>% 
    filter(PIS %in% PIS_matched) %>% 
    left_join(df_matched, by = c("PIS", "cbo")) %>% 
    mutate(observations = NA_integer_)
  
  saveRDS(df_clean, 
          paste0(personal_directory, 
                 "list_cbo_match_wages_public_past_years_df/cbo_match_wages_public_past_years_df", 
                 years[i+3], ".RDS"))
}


# 9) fazer uma lista dos CBOs que não foram classificados
files <- list.files(path = paste0(personal_directory, "list_cbo_n_match_wages_public_past_years/"), 
                    all.files = FALSE,
                    full.names = TRUE)

files_matched <- list.files(path = paste0(personal_directory, "list_cbo_match_wages_public_past_years/"), 
                            all.files = FALSE,
                            full.names = TRUE)

list_cbo_not_matched <- list()
cbo_matched <- c()
for(i in seq_along(files)){
  cbo_aux <- readRDS(files[i]) %>% distinct(cbo_original) %>% pull(cbo_original)
  
  cbo_matched <- c(cbo_matched, cbo_aux)
  
  list_cbo_not_matched[[i]] <- readRDS(files[i]) 
}

cbo_matched <- unique(cbo_matched)

cbo_no_match <- vec_cbo_no_direct_match[vec_cbo_no_direct_match %notin% cbo_matched]
# only the CBO 19245 (Arqueólogo) was not matched

cbo_n_matched_all <- bind_rows(list_cbo_not_matched)

# 10) para os PIS com match, tirar a média salarial por CBO e inputar essa média nos PIS sem match
files <- list.files(path = paste0(personal_directory, "list_cbo_n_match_wages_public_past_years/"), 
                    all.files = FALSE,
                    full.names = TRUE)

files_matched <- list.files(path = paste0(personal_directory, "list_cbo_match_wages_public_past_years/"), 
                            all.files = FALSE,
                            full.names = TRUE)

files_clean <- list.files(path = paste0(personal_directory, "list_cbo_no_direct_match_wages_federal/"), 
                          all.files = FALSE,
                          full.names = TRUE)

files_clean <- files_clean[4:length(files_clean)]

for (i in seq_along(files_matched)) {
  df_matched <- readRDS(files_matched[i]) %>% 
    group_by(cbo_original) %>% 
    summarise(wage_mean_mean = mean(as.numeric(wage_mean), na.rm = TRUE),
              wage_mean_median = median(as.numeric(wage_mean), na.rm = TRUE),
              wage_mean_min = min(as.numeric(wage_mean), na.rm = TRUE),
              wage_mean_max = max(as.numeric(wage_mean), na.rm = TRUE),
              wage_hour_mean = mean(as.numeric(wage_hour), na.rm = TRUE)) %>% 
    rename(cbo = cbo_original)
  
  PIS_n_matched <- readRDS(files[i]) %>% 
    distinct(PIS) %>% pull(PIS)
  
  df_clean <- readRDS(files_clean[i]) %>% 
    filter(PIS %in% PIS_n_matched) %>% 
    left_join(df_matched, by = "cbo") %>% 
    mutate(observations = NA_integer_)
  
  saveRDS(df_clean, 
          paste0(personal_directory, 
                 "list_cbo_n_match_wages_public_past_years_mean_cbo/cbo_n_match_wages_public_past_years_mean_cbo", 
                 years[i+3], ".RDS"))
}

# 11) juntando todas as bases finais
# direct match
files_direct_match <- list.files(path = paste0(personal_directory, "list_cbo_direct_match_wages_federal/"), 
                                 all.files = FALSE,
                                 full.names = TRUE)[4:10]
# direct past position
files_past_positions <- list.files(path = paste0(personal_directory, "list_cbo_match_wages_public_past_years_df/"), 
                            all.files = FALSE,
                            full.names = TRUE)

# mean wage for last positions with the same "original" CBO
files_mean_cbo_past_positions <- list.files(path = paste0(personal_directory, 
                                                          "list_cbo_n_match_wages_public_past_years_mean_cbo/"), 
                                            all.files = FALSE,
                                            full.names = TRUE)

# unindo as bases
for (i in seq_along(files_direct_match)) {
  df_direct_match <- readRDS(files_direct_match[i]) %>% 
    mutate(type = "Direct match by CBO")
  
  df_past_positions <- readRDS(files_past_positions[i]) %>% 
    select(-hours_contract_mean) %>% 
    mutate(type = "PIS past position in the private sector")
  
  df_mean_cbo_past_positions <- readRDS(files_mean_cbo_past_positions[i]) %>% 
    mutate(type = "Same CBO PIS past position in the private sector")
  
  df <- rbind(df_direct_match, df_past_positions, df_mean_cbo_past_positions)
  
  saveRDS(df, 
          paste0(personal_directory, 
                 "list_final_df/final_df_", 
                 years[i+3], ".RDS"))
  
  print(i)
}


files_final <- list.files(path = paste0(personal_directory, "list_final_df/"), 
                          all.files = FALSE,
                          full.names = TRUE)

df_final <- readRDS(files_final[4])

summary(df_final$wage_mean)
summary(df_final$wage_mean_mean)

summary(df_final$wage_hour)
summary(df_final$wage_hour_mean)

df_final %>% 
  select(PIS, wage_mean, wage_mean_mean) %>% 
  filter(wage_mean > 0 & wage_mean_mean > 0 & wage_mean < 10000 & wage_mean_mean < 10000) %>% 
  pivot_longer(cols = c("wage_mean", "wage_mean_mean"), names_to = "wage_type", values_to = "value") %>% 
  ggplot(aes(x = wage_type, y = value))+
  geom_violin() +
  # geom_boxplot(width = 0.1) +
  scale_color_grey() +
  theme_classic()

ggsave(filename = paste0(directory, "matching/dec2021/figures/vplot_wage_mean_vs_wage_mean_mean.png"),
       units = "cm", 
       width = 17,
       height = 10)

df_final %>% 
  select(PIS, wage_mean, wage_mean_mean, type) %>% 
  filter(wage_mean > 0 & wage_mean_mean > 0 & wage_mean < 10000 & wage_mean_mean < 10000) %>% 
  pivot_longer(cols = c("wage_mean", "wage_mean_mean"), names_to = "wage_type", values_to = "value") %>% 
  ggplot(aes(x = wage_type, y = value, color = type))+
  geom_violin() +
  # geom_boxplot(width = 0.1) +
  #scale_color_grey() +
  theme_classic()+
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.text = element_text(size = 6))

ggsave(filename = paste0(directory, "matching/dec2021/figures/vplot_wage_mean_vs_wage_mean_mean_by_type.png"),
       units = "cm", 
       width = 17,
       height = 10)

df_final %>% 
  select(PIS, wage_hour, wage_hour_mean) %>% 
  filter(wage_hour > 0 & wage_hour_mean > 0 & wage_hour < 500 & wage_hour_mean < 500) %>% 
  pivot_longer(cols = c("wage_hour", "wage_hour_mean"), names_to = "wage_type", values_to = "value") %>% 
  ggplot(aes(x = wage_type, y = value))+
  geom_violin() +
  # geom_boxplot(width = 0.1) +
  scale_color_grey() +
  theme_classic()

ggsave(filename = paste0(directory, "matching/dec2021/figures/vplot_wage_hour_vs_wage_hour_mean.png"),
       units = "cm", 
       width = 17,
       height = 10)

df_final %>% 
  select(PIS, wage_hour, wage_hour_mean, type) %>% 
  filter(wage_hour > 0 & wage_hour_mean > 0 & wage_hour < 500 & wage_hour_mean < 500) %>% 
  pivot_longer(cols = c("wage_hour", "wage_hour_mean"), names_to = "wage_type", values_to = "value") %>% 
  ggplot(aes(x = wage_type, y = value, color = type))+
  geom_violin() +
  # geom_boxplot(width = 0.1) +
  #scale_color_grey() +
  theme_classic()+
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.text = element_text(size = 6))

ggsave(filename = paste0(directory, "matching/dec2021/figures/vplot_wage_hour_vs_wage_hour_mean_by_type.png"),
       units = "cm", 
       width = 17,
       height = 10)

df_final_export <- df_final %>% 
  sample_n(250000) %>% 
  select(-PIS, -wage_dec, -wage_contract, -wage_last)

write.csv(df_final_export, file = paste0(personal_directory, "list_final_df/df_final_2001_sample.csv"))

```

## Commenting script step by step

### Years

Lula first term started in 2003, so the election took place in 2002. We wanted to evaluate the turnover two year before (2001 and 2002) and two years after Lula took office (2003 and 2004).

Besides that, since we decided to look at individual's working history to compute the cost of opportunity, we had to expand the search for the period from 1995 to 2004.

```{r year, eval = FALSE}
#years
years_lula <- as.character(c(2001:2004))
years <- as.character(c(1995:2004))
```

### Classifying occupations in public and private

Only a few occupations are restrict to public or private sector. In order to understand if a specific observation of RAIS (*individual-occupation*) is from the private or public sector, we have to look at the legal nature of the company where that occupation takes place.

In Brazil there are specific codes for public companies, separating all three federal levels: central/federal government, state government, and local government.

We are most interested in the central government' level's companies/firms: **c(1015, 1040, 1074, 1104, 1139, 1163, 1252, 10)**.

```{r nat_jur}
#http://www.mtecbo.gov.br/
nat_jur_fed <- as.character(c(1015, 1040, 1074, 1104, 1139, 1163, 1252, 10))
nat_jur <- as.character(c(10,
                          # Natureza Jurídica 1995
                          # https://concla.ibge.gov.br/estrutura/natjur-estrutura/natureza-juridica-1995.html
                          1015, #Poder Executivo Federal
                          1023, #Poder Executivo Estadual
                          1031, #Poder Executivo Municipal
                          1040, #Poder Legislativo Federal
                          1058, #Poder Legislativo Estadual
                          1066, #Poder Legislativo Municipal
                          1074, #Poder Judiciario Federal
                          1082, #Poder Judiciario Estadual
                          1090, #Orgão Autonomo de Direito Público
                          1104, #Autarquia Federal
                          1112, #Autarquia Estadual
                          1120, #Autarquia Municipal
                          1139, #Fundação Federal
                          1147, #Fundação Estadual
                          1155, #Fundação Municipal
                          1996, #Outras Formas de Organização da Administração Pública
                          2011, #Empresa Publica - Sociedade por Quotas de Responsabilidade Limitada
                          2020, #Empresa Publica - Sociedade Anonima de Capital Fechado
                          2038, #Sociedade Anonima de Capital Aberto com Controle Acionário Estatal (sociedade de economia mista) 
                          
                          # Natureza jurídica 2002
                          # https://concla.ibge.gov.br/estrutura/natjur-estrutura/natureza-juridica-2002.html
                          1015, #Órgão Público do Poder Executivo Federal
                          1023, #Órgão Público do Poder Executivo Estadual ou do Distrito Federal
                          1031, #Órgão Público do Poder Executivo Municipal
                          1040, #Órgão Público do Poder Legislativo Federal
                          1058, #Órgão Público do Poder Legislativo Estadual ou do Distrito Federal
                          1066, #Órgão Público do Poder Legislativo Municipal
                          1074, #Órgão Público do Poder Judiciário Federal
                          1082, #Órgão Público do Poder Judiciário Estadual
                          1104, #Autarquia Federal
                          1112, #Autarquia Estadual ou do Distrito Federal
                          1120, #Autarquia Municipal
                          1139, #Fundação Federal
                          1147, #Fundação Estadual ou do Distrito Federal
                          1155, #Fundação Municipal
                          1163, #Órgão Público Autônomo da União
                          1171, #Órgão Público Autônomo Estadual ou do Distrito Federal
                          1180, #Órgão Público Autônomo Municipal
                          2011, #Empresa Pública
                          2038)) #Sociedade de Economia Mista 

# There were changes in Natureza Jurídica after 2002, but none of them altering the classification of public entities.
# https://concla.ibge.gov.br/classificacoes/por-tema/organizacao-juridica
```

### Function to open RAIS files

We are working with the period going from from 1995 to 2004, and unfortunately columns names are not standardized among those years. I defined the necessary columns, look for their names into all df and add it to the function to have standardized df for the rest of the script.

The columns used in the script are the following:

-   "OCUPACAO": occupation code
-   "PIS": individual identifier
-   "NATUR JUR": company legal nature
-   "HORAS CONTR": number of hours the person works
-   "REM DEZ (R\$)": payment the person received in December
-   "REM MED (R\$)": average payment the person received in that year
-   "ULT REM": last payment the person received
-   "SAL CONTR": contractual wage
    -   the contractual wage might be different than the payment the person receives because of many reasons, such as benefits, payment for vacations or 13rd payment.

As it can be seem in the script, the first version of this script was importing more columns from the original df - what made everything slower and heavier/harder to process.

Since we decided to make the matching considering only the occupation code, comparing the same occupation in the public and private sector, I stopped importing individuals' characteristics.

```{r open RAIS matching}
open_files <- function(years_files = NA_character_, file_to_read = NA_character_){
  if(years_files == "2002"){
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("OCUPACAO", "PIS", "NATUR JUR",
                           "HORAS CONTR", "REM DEZ (R$)", "REM MED (R$)", "ULT REM", "SAL CONTR"
                           # ,"DT ADMISSAO", "RADIC CNPJ", "GRAU INSTR", "MES DESLIG", "CAUSA DESLI", "MUNICIPIO",
                           # "SEXO", "TEMP EMPR", "TP VINCL", "EMP EM 31/12", "DT NASCIMENT"
                           )) %>% 
      rename(natjur = `NATUR JUR`, 
             cbo = `OCUPACAO`,
             # date_adm = `DT ADMISSAO`,
             # CNPJ_root = `RADIC CNPJ`,
             # educ = `GRAU INSTR`,
             # month_quit = `MES DESLIG`,
             # reason_quit = `CAUSA DESLI`,
             # municipality = MUNICIPIO,
             # gender = SEXO,
             # time_employed = `TEMP EMPR`,
             # tp_contract = `TP VINCL`,
             # employed_dec = `EMP EM 31/12`,
             # b_date = `DT NASCIMENT`,
             hours_contract = `HORAS CONTR`,
             wage_dec = `REM DEZ (R$)`,
             wage_mean = `REM MED (R$)`,
             wage_last = `ULT REM`,
             wage_contract = `SAL CONTR`) %>% 
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             # age = as.numeric(years_files) - as.numeric(substr(b_date,(nchar(b_date)+1)-4,nchar(b_date))),
             # date_adm = ifelse(nchar(date_adm) == 7, paste0("0", date_adm), date_adm),
             # date_adm = as.Date(date_adm, format = "%d%m%Y"),
             # date_quit = ifelse(month_quit == "00", NA, 
             #                    as.Date(paste0("30/", month_quit, "/", years_files), format = "%d/%m/%Y")),
             # wage_dec = as.numeric(str_replace(wage_dec, ",", ".")),
             wage_mean = as.numeric(str_replace(wage_mean, ",", ".")),
             wage_last = as.numeric(str_replace(wage_last, ",", ".")),
             wage_contract = as.numeric(str_replace(wage_contract, ",", ".")),
             wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                                as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) #%>% 
      # select(-b_date, -month_quit)
      # filter(cbo %in% c("20496", "20512", "20528", "20544",
      #                     "21420", "21490", 
      #                     "24592", "24608", "24624", "24640", 
      #                     "16416", "16432", "16448", 
      #                     "28688", "28704", 
      #                     "39385"))

  } else if (years_files %in% as.character(1995:1998)){
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("CBO 94 Ocupação", "PIS", "Natureza Jurídica", 
                           "Vl Remun Dezembro (SM)", "Vl Remun Média (SM)", "Qtd Hora Contr"
                           # ,"Ano Admissão", "CNPJ Raiz", "Grau Instrução 2005-1985", "Mês Admissão",
                           # "Mês Desligamento", "Motivo Desligamento", "Município", "Sexo Trabalhador", "Tempo Emprego",
                           # "Tipo Vínculo", "Vínculo Ativo 31/12", "Idade", "Tipo Admissão"
                )) %>% 
      rename(natjur = `Natureza Jurídica`, 
             cbo = `CBO 94 Ocupação`,
             # year_adm = `Ano Admissão`,
             # CNPJ_root = `CNPJ Raiz`,
             # educ = `Grau Instrução 2005-1985`,
             # month_adm = `Mês Admissão`,
             # month_quit = `Mês Desligamento`,
             # reason_quit = `Motivo Desligamento`,
             # municipality = `Município`,
             # gender = `Sexo Trabalhador`,
             # time_employed = `Tempo Emprego`,
             # tp_contract = `Tipo Vínculo`,
             # employed_dec = `Vínculo Ativo 31/12`,
             # age = `Idade`,
             hours_contract = `Qtd Hora Contr`,
             wage_dec = `Vl Remun Dezembro (SM)`,
             wage_mean = `Vl Remun Média (SM)`) %>% 
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             wage_last = NA_character_,
             wage_contract = NA_character_,
             wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                                as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) # %>% 
    # select(-month_adm, -year_adm, -month_quit)
    # filter(cbo %in% c("20496", "20512", "20528", "20544",
    #                     "21420", "21490", 
    #                     "24592", "24608", "24624", "24640", 
    #                     "16416", "16432", "16448", 
    #                     "28688", "28704", 
    #                     "39385"))
    
  } else if (years_files %in% as.character(1995)){
    df <- df %>% 
      mutate(# Maio 1995,  Lei nº 9032, de 1995, salário mínimo: R$ 100,00
        wage_dec = as.numeric(str_replace(wage_dec, ",", ".")) * 100,
        wage_mean = as.numeric(str_replace(wage_mean, ",", ".")) * 100,
        wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                           as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) 
  } else if (years_files %in% as.character(1996)){
    df <- df %>% 
      mutate(# Maio 1996,  Lei nº 9971, de 1996, salário mínimo: R$ 120,00
        wage_dec = as.numeric(str_replace(wage_dec, ",", ".")) * 112,
        wage_mean = as.numeric(str_replace(wage_mean, ",", ".")) * 112,
        wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                           as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) 
  } else if (years_files %in% as.character(1997)){
    df <- df %>% 
      mutate(# Maio 1997,  Lei nº 9971, de 2000, salário mínimo: R$ 130,00
        wage_dec = as.numeric(str_replace(wage_dec, ",", ".")) * 120,
        wage_mean = as.numeric(str_replace(wage_mean, ",", ".")) * 120,
        wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                           as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) 
  } else if (years_files %in% as.character(1998)){
    df <- df %>% 
      mutate(# Maio 1998,  Lei nº 9971, de 2000, salário mínimo: R$ 130,00
             wage_dec = as.numeric(str_replace(wage_dec, ",", ".")) * 130,
             wage_mean = as.numeric(str_replace(wage_mean, ",", ".")) * 130,
             wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                                as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) 
  } else if (years_files %in% as.character(c(1999:2001))){
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("CBO 94 Ocupação", "PIS", "Natureza Jurídica", 
                           "Vl Remun Dezembro Nom", "Vl Remun Média Nom", "Qtd Hora Contr"
                           # ,"Ano Admissão", "CNPJ Raiz", "Grau Instrução 2005-1985", "Mês Admissão",
                           # "Mês Desligamento", "Motivo Desligamento", "Município", "Sexo Trabalhador", "Tempo Emprego",
                           # "Tipo Vínculo", "Vínculo Ativo 31/12", "Idade", "Tipo Admissão"
                           )) %>% 
      rename(natjur = `Natureza Jurídica`, 
             cbo = `CBO 94 Ocupação`,
             # year_adm = `Ano Admissão`,
             # CNPJ_root = `CNPJ Raiz`,
             # educ = `Grau Instrução 2005-1985`,
             # month_adm = `Mês Admissão`,
             # month_quit = `Mês Desligamento`,
             # reason_quit = `Motivo Desligamento`,
             # municipality = `Município`,
             # gender = `Sexo Trabalhador`,
             # time_employed = `Tempo Emprego`,
             # tp_contract = `Tipo Vínculo`,
             # employed_dec = `Vínculo Ativo 31/12`,
             # age = `Idade`,
             hours_contract = `Qtd Hora Contr`,
             wage_dec = `Vl Remun Dezembro Nom`,
             wage_mean = `Vl Remun Média Nom`)%>% 
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             # date_adm = as.Date(paste0("01/", month_adm, "/", year_adm), format = "%d/%m/%Y"),
             # date_quit = ifelse(month_quit == "00", NA, 
             #                    as.Date(paste0("30/", month_quit, "/", years_files), format = "%d/%m/%Y")),
             wage_dec = as.numeric(str_replace(wage_dec, ",", ".")),
             wage_mean = as.numeric(str_replace(wage_mean, ",", ".")),
             wage_last = NA_character_,
             wage_contract = NA_character_,
             wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                                as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) #%>% 
      # select(-month_adm, -year_adm, -month_quit)
      # filter(cbo %in% c("20496", "20512", "20528", "20544",
      #                     "21420", "21490", 
      #                     "24592", "24608", "24624", "24640", 
      #                     "16416", "16432", "16448", 
      #                     "28688", "28704", 
      #                     "39385"))
    
  } else{
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("OCUPACAO 94", "PIS", "NATUR JUR", 
                           "HORAS CONTR", "REM DEZ (R$)", "REM MED (R$)", "ULT REM", "SAL CONTR"
                           # ,"DT ADMISSAO", "RADIC CNPJ", "GRAU INSTR", "MES DESLIG", "CAUSA DESLI", 
                           # "MUNICIPIO", "SEXO", "TEMP EMPR", "TP VINCULO", "EMP EM 31/12", "DT NASCIMENT"
                           )) %>% 
      rename(natjur = `NATUR JUR`, 
             cbo = `OCUPACAO 94`,
             # date_adm = `DT ADMISSAO`,
             # CNPJ_root = `RADIC CNPJ`,
             # educ = `GRAU INSTR`,
             # month_quit = `MES DESLIG`,
             # reason_quit = `CAUSA DESLI`,
             # municipality = MUNICIPIO,
             # gender = SEXO,
             # time_employed = `TEMP EMPR`,
             # tp_contract = `TP VINCULO`,
             # employed_dec = `EMP EM 31/12`,
             # b_date = `DT NASCIMENT`,
             hours_contract = `HORAS CONTR`,
             wage_dec = `REM DEZ (R$)`,
             wage_mean = `REM MED (R$)`,
             wage_last = `ULT REM`,
             wage_contract = `SAL CONTR`) %>% 
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             # age = as.numeric(years_files) - as.numeric(substr(b_date,(nchar(b_date)+1)-4,nchar(b_date))),
             # date_adm = ifelse(nchar(date_adm) == 7, paste0("0", date_adm), date_adm),
             # date_adm = as.Date(date_adm, format = "%d%m%Y"),
             # date_quit = ifelse(month_quit == "00", NA, 
             #                    as.Date(paste0("30/", month_quit, "/", years_files), format = "%d/%m/%Y")),
             wage_dec = as.numeric(str_replace(wage_dec, ",", ".")),
             wage_mean = as.numeric(str_replace(wage_mean, ",", ".")),
             wage_last = as.numeric(str_replace(wage_last, ",", ".")),
             wage_contract = as.numeric(str_replace(wage_contract, ",", ".")),
             wage_hour = ifelse(as.numeric(hours_contract) == 0, 0, 
                                as.numeric(str_replace(wage_mean, ",", "."))/as.numeric(hours_contract))) #%>% 
      # select(-b_date, -month_quit)
    # filter(cbo %in% c("20496", "20512", "20528", "20544",
    #                       "21420", "21490", 
    #                       "24592", "24608", "24624", "24640", 
    #                       "16416", "16432", "16448", 
    #                       "28688", "28704", 
    #                       "39385"))
  }
  
  return(df)
}

```

### Listing and classifying CBOs

For each year, I list all CBO-positions and classify them as belonging to the private or public sector. Then I save the data.

I make a list of unique public and federal CBOs for each year, bind the list of all years in one df and keep unique CBOs only.

```{r list classify CBO}
list_cbo <- list()
for(i in seq_along(years)){
  
  print(years[i])
  
  # list files for each year
  files <- grep(list.files(path = paste0(directory, years[i]), 
                           all.files = FALSE,
                           full.names = TRUE), 
                pattern= ('Estb|ESTB|IGNORADO'), 
                inv=TRUE, value=TRUE)
  
  list_data_temp <- list()
  list_cbo_temp <- list()
  for(j in seq_along(files)) { 
    # 1) Para os anos de 2001, 2002, 2003 e 2004 (2 antes e 2 depois de Lula) classificar todas as 
    # observações em público ou privado com base em TODAS as Naturezas Jurídicas;
    list_data_temp[[j]] <- open_files(years_file = years[i], file_to_read = files[j]) %>% 
      mutate(public = 0, public = ifelse(natjur %in% nat_jur, 1, public),
             federal = 0, federal = ifelse(natjur %in% nat_jur_fed, 1, federal)) %>% 
    # 2) Para a base de dados classificada, filtrar para CBO privado OU público E federal;
      filter(public == 0 | federal == 1) %>% 
    # criar marcação de ano
      mutate(year = years[i])
    
    # 3) Gerar uma lista com CBO públicos federais para esses anos;
    list_cbo_temp[[j]] <- list_data_temp[[j]] %>% 
      filter(federal == 1) %>% 
      distinct(cbo) %>% 
      select(cbo)
    
    print(j)
  }
  
  df <- bind_rows(list_data_temp)
  saveRDS(df, paste0(personal_directory, "list_data_wages/data_wages_", years[i], ".RDS"))
  
  list_cbo[[i]] <- bind_rows(list_cbo_temp) %>% distinct(cbo)
  
  remove(df, list_data_temp, list_cbo_temp)
}

# 4) Manualmente classificar os CBOs públicos federais que NÃO PODEM ser encontrados no setor público (ex. juiz);
cbo_federais <- bind_rows(list_cbo) %>% distinct(cbo)

```

### Manual cleaning of CBO classification

since RAIS is a register filled by the company/employer, it might contain errors. Looking at occupational codes, it is clear that some of them can not belong to the private sector. You can see some examples in the script bellow.

If I find those codes in the private sector, I can infer that there was a mistake filling the RAIS forms. Since I can not correct this mistake (I don't know the real occupation code, of if the mistake came from the legal nature code), I decided to exclude those observation (*this was something Paula and I discussed*).

**EXCLUSION:** observations that were classified as private (given its legal nature code), but had a occupation code that can only be public.

**NUMBER OF OBSERVATIONS:** initial number of observations - final number of observations = number of observations deleted.

As we can see in the table, a very small proportion of observations were deleted. The matching used the observations from the "Final nº of observations" column.

| Year | Initial nº of obs | Final nº of obs | Deleted |
|:----:|:-----------------:|:---------------:|:-------:|
| 1995 |    27,562,711     |   27,541,093    | 21,618  |
| 1996 |    26,516,841     |   26,493,344    | 23,497  |
| 1997 |    27,863,657     |   27,837,125    | 26,532  |
| 1998 |    28,194,129     |   28,158,249    | 35,880  |
| 1999 |    28,460,449     |   28,404,008    | 56,441  |
| 2000 |    30,473,813     |   30,428,651    | 45,162  |
| 2001 |    32,019,990     |   31,966,857    | 53,133  |
| 2002 |    36,986,619     |   36,914,780    | 71,839  |
| 2003 |    33,930,145     |   33,878,305    | 51,840  |
| 2004 |    36,376,481     |   36,325,087    | 51,394  |

```{r cleaning CBO classification}
# write.csv(cbo_federais, file = paste0(personal_directory, "cbos_federal.csv"))

remove(list_cbo)
# CBOs que são apenas públicos:
# No CBO 94 não havia presidente, por exemplo. Está dentro de dirigentes do setor público
# 31920 >	Agente de saúde pública
# 21420 >	Funcionário público federal superior (1114-05 no CBO 2002)
# 31490 >	Outros serventuários da justiça e trabalhadores assemelhados
# 31390 >	Outros agentes superiores de polícia
# 21490 >	Outros funcionários públicos superiores
# 21320 >	Juiz federal
# 31440 >	Oficial de justiça
# 21330 >	Juiz estadual
# 31430 >	Escrivão
# 58290 >	Outros policiais e trabalhadores assemelhados
# 31420 >	Tabelião
# 21390 >	Outros membros superiores do Poder Judiciário
# 21350 >	Juiz do trabalho
# 21340 >	Juiz militar
# 21220 >	Membro superior do Poder Executivo
# 21130 >	Deputado federal
# 21360 >	Juiz eleitoral
# 31320 >	Delegado de polícia
# 21120 >	Senador
# 21140 >	Vereador
# 21430 >	Funcionário público estadual superior
# 12920 >	Procurador da fazenda nacional
# 58250 >	Papiloscopista policial
# 31340 >	Perito criminal
# 58430 >	Policial rodoviário
# 39385 >	Despachante aduaneiro
# 58230 >	Detetive de polícia
cbo_only_public <- as.character(c(31920, 21420, 31490, 31390, 21490, 21320, 31440, 21330, 31430, 
                                  58290, 31420, 21390, 21350, 21340, 21220, 21130, 21360, 31320, 
                                  21120, 21140, 21430, 12920, 58250, 31340, 58430, 39385, 58230))

# 5) Excluir da base CBO que são públics e foram classificados como privado. 
# Decidimos excluir porque não sabemos se o erro foi no preenchimento do CBO ou da Natureza Jurídica;
# list_data <- readRDS(paste0(personal_directory, "list_data.RDS"))

files <- list.files(path = paste0(personal_directory, "list_data_wages/"), 
                    all.files = FALSE,
                    full.names = TRUE)

for (i in seq_along(files)) {
  df <- readRDS(files[i])
  
  print(years[i])
  print(nrow(df))
  # os CBOs que não estão na lista podem ser públicos e privados
  df1 <- df %>% filter(cbo %notin% cbo_only_public)
  
  # para os CBOs que estão na lista só vamos manter as observações corretamente classificadas como públicas
  df2 <- df %>% 
    filter(cbo %in% cbo_only_public) %>% 
    filter(federal == 1)
  
  df <- rbind(df1, df2)
  print(nrow(df))
  
  saveRDS(df, paste0(personal_directory, "/list_data_wages_clean/data_wages_clean_", years[i], ".RDS"))
  
  remove(df, df1, df2)
}

# Número de observações que perdemos por ano:
# 1995: 27,562,711 - 27,541,093 = 21,618
# 1996: 26,516,841 - 26,493,344 = 23,497
# 1997: 27,863,657 - 27,837,125 = 26,532
# 1998: 28,194,129 - 28,158,249 = 35,880
# 1999: 28,460,449 - 28,404,008 = 56,441
# 2000: 30,473,813 - 30,428,651 = 45,162
# 2001: 32,019,990 - 31,966,857 = 53,133
# 2002: 36,986,619 - 36,914,780 = 71,839
# 2003: 33,930,145 - 33,878,305 = 51,840
# 2004: 36,376,481 - 36,325,087 = 51,394
```

### Matching

#### Withing the same year.

The matching was made at occupation level. First within the same year, and after looking for previous years.

One occupation was considered to have a matching if it appears, in the same year, both in the private and in the public sector. For those, I computed average statistics for wage and hour of contract in the public and in the private sector (save in the df `df_cbo_matched`).

I also saved the occupations that did not have a match in that year. For those, I will look for a matching in the previous years.

```{r}
# list_data <- readRDS(paste0(personal_directory, "list_data_wages_clean.RDS"))
# 6) Com essa base limpa, fazer o match ano a ano. 
files <- list.files(path = paste0(personal_directory, "list_data_wages_clean/"), 
                    all.files = FALSE,
                    full.names = TRUE)

list_cbo_matched <- list()
list_cbo_not_matched <- list()
for (i in seq_along(files)) {
  print(years[i])
  
  df <- readRDS(files[i])
  
  cbos_duplicates <- df %>% 
    count(cbo, public) %>% 
    add_count(cbo, name = "duplicates") %>% 
    filter(duplicates > 1) %>% 
    ungroup() %>% 
    distinct(cbo) %>% 
    pull(cbo)
  
  df_cbo_matched <- df %>% 
    filter(cbo %in% cbos_duplicates) %>% 
    group_by(cbo, public) %>% 
    mutate(observations = n(),
           wage_mean_mean = mean(as.numeric(wage_mean), na.RM = TRUE),
           wage_mean_median = median(as.numeric(wage_mean), na.RM = TRUE),
           # wage_mean_var = ifelse(n %in% c(1, 2), 0, var(as.numeric(wage_mean), na.RM = TRUE)),
           wage_mean_min = min(as.numeric(wage_mean), na.RM = TRUE),
           wage_mean_max = max(as.numeric(wage_mean), na.RM = TRUE),
           hours_contract = mean(as.numeric(hours_contract), na.rm = TRUE),
           wage_hour_mean = mean(as.numeric(wage_hour), na.rm = TRUE),
           year = years[i]) %>% 
    ungroup() 
  
  saveRDS(df_cbo_matched, 
          paste0(personal_directory, "list_cbo_direct_match_wages/cbo_direct_match_wages_", years[i], ".RDS"))
  
  df_cbo_no_matched <- df %>% 
    filter(cbo %notin% cbos_duplicates) %>% 
    filter(public == 1) %>% 
    mutate(year = years[i]) %>% 
    ungroup()
  
  saveRDS(df_cbo_no_matched, 
          paste0(personal_directory, "list_cbo_no_direct_match_wages/cbo_no_direct_match_wages_", years[i], ".RDS"))
  
  remove(df, df_cbo_matched, df_cbo_no_matched)
    
}

files <- list.files(path = paste0(personal_directory, "list_cbo_direct_match_wages/"), 
                    all.files = FALSE,
                    full.names = TRUE)

for(i in seq_along(files)){
  df <- readRDS(files[i])
  
  df <- df %>% 
    filter(federal == 1)
  
  saveRDS(df, 
          paste0(personal_directory, "list_cbo_direct_match_wages_federal/cbo_direct_match_wages_federal_", 
                 years[i], ".RDS"))
}
```

#### Previous years

First I make a vector of all CBO (ocuppations) that were federal and did not had a matching in that specific year.

```{r}
# 7) Fazer uma lista dos CBOs classificados como público que não tiveram match;
files <- list.files(path = paste0(personal_directory, "list_cbo_no_direct_match_wages/"), 
                    all.files = FALSE,
                    full.names = TRUE)

vec_cbo_no_direct_match <- c()
for(i in seq_along(files)){
  df <- readRDS(files[i])
  
  df <- df %>% 
    filter(federal == 1)
  
  vec_cbo_no_direct_match <- c(vec_cbo_no_direct_match, 
                               (df %>% distinct(cbo) %>% pull(cbo)))
  
  saveRDS(df, 
          paste0(personal_directory, "list_cbo_no_direct_match_wages_federal/cbo_no_direct_match_wages_federal_", 
                 years[i], ".RDS"))
}

vec_cbo_no_direct_match <- unique(vec_cbo_no_direct_match)
```

To look in the past, I need both occupational codes and individuals' IDs. For those people in occupations in the federal public sector that did not had a matching in one specific year, I want to find the last occupations they had in the last three years in the private sector.
I will then match the actual occupation (CBO from public sector that did not had a match in the private sector in that specific year) with the previous occupation (CBO from the last position the person had in the private sector in a 3 years window).
I excluded from the search PIS that are cleary wrong,  and could affect the matching (`PIS %notin% c("00000000000", "0", "99999999999")`)
To make the wages comparable, I correct the previous wages by INPC, an inflation index for wages in BRazil.

<div class = "orange">
**PARA PAULA**: para mim é um pouco complicado de explicar, então escrevo aqui também em portugues.
Primeiro eu olhei dentro do ano, e peguei todos os CBOs que apareciam tanto no privado quanto no público.
Alguns CBOs do setor público não apareciam no setor privado. Eu guardo esses para busca-los no passado.
Mas não posso buscar CBOs no passado. Então para esses CBOs sem match, eu pego os PIS das pessoas que estavam trabalhando nessas ocupações.
Daí eu uso esse PIS para buscar no passado. Eu olho se esse PIS aparece na RAIS nos 3 anos anteriores ao ano de referência em alguma posição no setor privado (porque no ano base a pessoa/PIS está emrpegada no setor público).
Se sim, então eu pego o CBO dessa ocupação. 
Assim, para cada pessoa eu tenho o CBO do ano base (ocupação atual no setor público) e o CBO do último emprego privado nos 3 anos anteriores (CBO do passado no setor privado).
O match se dá entre esses dois CBOs.
Tudo com correção monetário pelo INPC.

Para as pessoas que eu não acho no passado, eu uso a média do opportunity cost daquele mesmo CBO, considerando as pessoas que sim eu acho no passado.
</div>



```{r}
# 8) Para os CBOs de 7), buscar outras posições das pessoas nos três anos anteriores (menor ano será 1995) no setor privado.
files <- list.files(path = paste0(personal_directory, "list_cbo_no_direct_match_wages_federal/"), 
                    all.files = FALSE,
                    full.names = TRUE)

files_bsearch <- list.files(path = paste0(personal_directory, "list_data_wages_clean/"), 
                    all.files = FALSE,
                    full.names = TRUE)

rename_fun <- function(x){
  x <- paste0(x, "2")
  
  return(x)
}

sub_comma_fun <- function(str){
  str <- gsub(",", ".", str)
  
  return(str)
}

inflation_index <- read_csv(paste0(personal_directory, "input/INPC_correction.csv"), 
                            col_types = cols(INPC = col_number(), 
                                             year_destination = col_character(), 
                                             year_origin = col_character()))
for(i in length(files):4){
  df <- readRDS(files[i])
  
  CBO_vector <- df %>% 
    distinct(cbo) %>% 
    pull(cbo)
  
  df <- df %>% 
    mutate_at(c("hours_contract", "wage_dec", "wage_mean", "wage_last", "wage_contract", 
                "public", "federal", "wage_hour"), as.numeric) %>% 
    mutate_at(c("cbo", "PIS", "natjur", "year"), as.character) %>% 
    filter(cbo %in% CBO_vector) %>% 
  # excluir PIS "errados"
    filter(PIS %notin% c("00000000000", "0", "99999999999"))
  
  PIS_vector <- df %>% 
    filter(PIS != "0") %>% 
    pull(PIS)

  list_aux <- list()
  PIS_matched <- c()
  j <- 1
  for (k in c(i - 1, i - 2, i - 3)) { #starts at year - 1
    if(k <= 0){
      print(k)
    } 
    else {
      df_bsearch <- readRDS(files_bsearch[k]) %>% 
        # excluir PIS "errados"
        filter(PIS %notin% c("00000000000", "0", "99999999999"))
      
      inflation <- inflation_index %>% 
        filter(year_destination == years[i],
               year_origin == years[k]) %>% 
        pull(INPC)
      
      list_aux[[j]] <- df_bsearch %>% 
        filter(public == 0) %>% 
        filter(PIS %in% PIS_vector) %>% 
        mutate_at(c("wage_dec", "wage_mean", "wage_last", "wage_contract"), sub_comma_fun) %>% 
        mutate_at(c("hours_contract", "wage_dec", "wage_mean", "wage_last", "wage_contract", 
                    "public", "federal", "wage_hour"), as.numeric) %>% 
        mutate_at(c("cbo", "PIS", "natjur", "year"), as.character) %>% 
        mutate(wage_dec = wage_dec*inflation, 
               wage_mean = wage_mean*inflation, 
               wage_last = wage_last*inflation, 
               wage_contract = wage_contract*inflation,
               wage_hour = wage_hour*inflation)
        
      PIS_matched <- c(PIS_matched, 
                       (list_aux[[j]] %>% pull(PIS)))
      
      PIS_vector <- PIS_vector[PIS_vector %notin% PIS_matched]
      
      j <- j + 1
      
      remove(df_bsearch)
    }
  }
  df2 <- bind_rows(list_aux) %>% 
    distinct(PIS, cbo, hours_contract, wage_contract, wage_hour, year, .keep_all = TRUE)
  
  df <- df %>% 
    select(PIS, cbo) %>% 
    rename(cbo_original = cbo) %>%
    left_join(df2, by = "PIS")
  
  df_matched <- df %>% 
    filter(!is.na(cbo)) %>% 
    add_count(cbo_original, cbo)
  
  saveRDS(df_matched, 
          paste0(personal_directory, 
                 "list_cbo_match_wages_public_past_years/cbo_match_wages_public_past_years_", 
                 years[i], ".RDS"))
  
  df_n_matched <- df %>% 
    filter(is.na(cbo))
  
  saveRDS(df_n_matched, 
          paste0(personal_directory, 
                 "list_cbo_n_match_wages_public_past_years/list_cbo_n_match_wages_public_past_years_", 
                 years[i], ".RDS"))
  
  print(i)
  
  remove(df, df_matched, df_n_matched)
}

```

After that, I standardize column names to combine this df with the "withing-year-match" df.
```{r}
# 8a) padronizar o banco de dados para juntar com os outros depois
files_matched <- list.files(path = paste0(personal_directory, "list_cbo_match_wages_public_past_years/"), 
                            all.files = FALSE,
                            full.names = TRUE)

files_clean <- list.files(path = paste0(personal_directory, "list_cbo_no_direct_match_wages_federal/"), 
                          all.files = FALSE,
                          full.names = TRUE)

files_clean <- files_clean[4:length(files_clean)]

for (i in seq_along(files_matched)) {
  df_matched <- readRDS(files_matched[i]) %>% 
    select(PIS, cbo_original, wage_mean, hours_contract, wage_hour) %>% 
    rename(wage_mean_mean = wage_mean,
           wage_hour_mean = wage_hour,
           hours_contract_mean = hours_contract,
           cbo = cbo_original) %>% 
    mutate(wage_mean_median = NA_integer_,
           wage_mean_min = NA_integer_,
           wage_mean_max = NA_integer_) 
  
  PIS_matched <- df_matched %>% 
    distinct(PIS) %>% pull(PIS)
  
  df_clean <- readRDS(files_clean[i]) %>% 
    filter(PIS %in% PIS_matched) %>% 
    left_join(df_matched, by = c("PIS", "cbo")) %>% 
    mutate(observations = NA_integer_)
  
  saveRDS(df_clean, 
          paste0(personal_directory, 
                 "list_cbo_match_wages_public_past_years_df/cbo_match_wages_public_past_years_df", 
                 years[i+3], ".RDS"))
}

```

Even after that there might by CBOs that did not had a matching. I made a list of them, and found out that the only CBO not matched was **19245 (Arqueólogo)**.  
```{r}
# 9) fazer uma lista dos CBOs que não foram classificados
files <- list.files(path = paste0(personal_directory, "list_cbo_n_match_wages_public_past_years/"), 
                    all.files = FALSE,
                    full.names = TRUE)

files_matched <- list.files(path = paste0(personal_directory, "list_cbo_match_wages_public_past_years/"), 
                            all.files = FALSE,
                            full.names = TRUE)

list_cbo_not_matched <- list()
cbo_matched <- c()
for(i in seq_along(files)){
  cbo_aux <- readRDS(files[i]) %>% distinct(cbo_original) %>% pull(cbo_original)
  
  cbo_matched <- c(cbo_matched, cbo_aux)
  
  list_cbo_not_matched[[i]] <- readRDS(files[i]) 
}

cbo_matched <- unique(cbo_matched)

cbo_no_match <- vec_cbo_no_direct_match[vec_cbo_no_direct_match %notin% cbo_matched]
# only the CBO 19245 (Arqueólogo) was not matched

cbo_n_matched_all <- bind_rows(list_cbo_not_matched)
```

For those PIS/CBO that had a match in the past, I included the median wage in the original df, and then combined all df in just one.
```{r}
# 10) para os PIS com match, tirar a média salarial por CBO e inputar essa média nos PIS sem match
files <- list.files(path = paste0(personal_directory, "list_cbo_n_match_wages_public_past_years/"), 
                    all.files = FALSE,
                    full.names = TRUE)

files_matched <- list.files(path = paste0(personal_directory, "list_cbo_match_wages_public_past_years/"), 
                            all.files = FALSE,
                            full.names = TRUE)

files_clean <- list.files(path = paste0(personal_directory, "list_cbo_no_direct_match_wages_federal/"), 
                          all.files = FALSE,
                          full.names = TRUE)

files_clean <- files_clean[4:length(files_clean)]

for (i in seq_along(files_matched)) {
  df_matched <- readRDS(files_matched[i]) %>% 
    group_by(cbo_original) %>% 
    summarise(wage_mean_mean = mean(as.numeric(wage_mean), na.rm = TRUE),
              wage_mean_median = median(as.numeric(wage_mean), na.rm = TRUE),
              wage_mean_min = min(as.numeric(wage_mean), na.rm = TRUE),
              wage_mean_max = max(as.numeric(wage_mean), na.rm = TRUE),
              wage_hour_mean = mean(as.numeric(wage_hour), na.rm = TRUE)) %>% 
    rename(cbo = cbo_original)
  
  PIS_n_matched <- readRDS(files[i]) %>% 
    distinct(PIS) %>% pull(PIS)
  
  df_clean <- readRDS(files_clean[i]) %>% 
    filter(PIS %in% PIS_n_matched) %>% 
    left_join(df_matched, by = "cbo") %>% 
    mutate(observations = NA_integer_)
  
  saveRDS(df_clean, 
          paste0(personal_directory, 
                 "list_cbo_n_match_wages_public_past_years_mean_cbo/cbo_n_match_wages_public_past_years_mean_cbo", 
                 years[i+3], ".RDS"))
}

# 11) juntando todas as bases finais
# direct match
files_direct_match <- list.files(path = paste0(personal_directory, "list_cbo_direct_match_wages_federal/"), 
                                 all.files = FALSE,
                                 full.names = TRUE)[4:10]
# direct past position
files_past_positions <- list.files(path = paste0(personal_directory, "list_cbo_match_wages_public_past_years_df/"), 
                            all.files = FALSE,
                            full.names = TRUE)

# mean wage for last positions with the same "original" CBO
files_mean_cbo_past_positions <- list.files(path = paste0(personal_directory, 
                                                          "list_cbo_n_match_wages_public_past_years_mean_cbo/"), 
                                            all.files = FALSE,
                                            full.names = TRUE)

# unindo as bases
for (i in seq_along(files_direct_match)) {
  df_direct_match <- readRDS(files_direct_match[i]) %>% 
    mutate(type = "Direct match by CBO")
  
  df_past_positions <- readRDS(files_past_positions[i]) %>% 
    select(-hours_contract_mean) %>% 
    mutate(type = "PIS past position in the private sector")
  
  df_mean_cbo_past_positions <- readRDS(files_mean_cbo_past_positions[i]) %>% 
    mutate(type = "Same CBO PIS past position in the private sector")
  
  df <- rbind(df_direct_match, df_past_positions, df_mean_cbo_past_positions)
  
  saveRDS(df, 
          paste0(personal_directory, 
                 "list_final_df/final_df_", 
                 years[i+3], ".RDS"))
  
  print(i)
}
```

## Conclusion
The matching is at CBO level. 

- If the CBO appears both in the public and private sector withing one year, the opportunity cost for those in the public sector will be the average wage of the same position in the same year in the private sector.
- If the CBO only appears in the public sector in that year, I will look what the people working in this position this year were doing in the private sector in the previous three years.
    - If the person appeared in the private sector in the previous three years, her opportunity cost will be her last wage, corrected by inflation;
    - If the person does not appears in the private sector in the previous three years, she will have an opportunity cost that is an average. This average considers what people with the same current CBO were earning in the previous years.



# **2a_matching_public_private_add_columns_v2.R**

Here is the full script You can hide it and go to the next section where I comment it step by step.

```{r full code, eval=FALSE}
#####################
#### DEFINITIONS ####
#####################
rm(list=ls())
gc()

library("tidyverse")
library("readxl")
library("data.table")

# Para facilitar
select <- dplyr::select
filter <- dplyr::filter
`%notin%` = Negate(`%in%`)

personal_directory <- "C:/Users/taina/Desktop/RAIS/matching/dec2021/"
directory <- "C:/Users/taina/Desktop/RAIS/"

years <- as.character(c(1998:2004))

##################################
#### FUNCTION TO OPEN RAIS DF ####
##################################
open_files <- function(years_files = NA_character_, file_to_read = NA_character_){
  if(years_files == "2002"){
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("OCUPACAO", "PIS", 
                           "DT ADMISSAO", "RADIC CNPJ", "GRAU INSTR", "MES DESLIG", "CAUSA DESLI", "MUNICIPIO",
                           "SEXO", "TEMP EMPR", "TP VINCL", "EMP EM 31/12", "DT NASCIMENT")) %>% 
      rename(cbo = `OCUPACAO`,
             date_adm = `DT ADMISSAO`,
             CNPJ_root = `RADIC CNPJ`,
             educ = `GRAU INSTR`,
             month_quit = `MES DESLIG`,
             reason_quit = `CAUSA DESLI`,
             municipality = MUNICIPIO,
             gender = SEXO,
             time_employed = `TEMP EMPR`,
             tp_contract = `TP VINCL`,
             employed_dec = `EMP EM 31/12`,
             b_date = `DT NASCIMENT`) %>%
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             age = as.numeric(years_files) - as.numeric(substr(b_date,(nchar(b_date)+1)-4,nchar(b_date))),
             date_adm = ifelse(nchar(date_adm) == 7, paste0("0", date_adm), date_adm),
             date_adm = as.Date(date_adm, format = "%d%m%Y"),
             date_quit = as.Date(paste0("30/", month_quit, "/", years_files), format = "%d/%m/%Y"),
             CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root)))) %>%
      select(-b_date, -month_quit)
    
  } else if (years_files %in% as.character(1995:1998)){
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("CBO 94 Ocupação", "PIS", 
                           "Ano Admissão", "CNPJ Raiz", "Grau Instrução 2005-1985", "Mês Admissão",
                           "Mês Desligamento", "Motivo Desligamento", "Município", "Sexo Trabalhador", "Tempo Emprego",
                           "Tipo Vínculo", "Vínculo Ativo 31/12", "Idade", "Tipo Admissão")) %>%
      rename(cbo = `CBO 94 Ocupação`,
             year_adm = `Ano Admissão`,
             CNPJ_root = `CNPJ Raiz`,
             educ = `Grau Instrução 2005-1985`,
             month_adm = `Mês Admissão`,
             month_quit = `Mês Desligamento`,
             reason_quit = `Motivo Desligamento`,
             municipality = `Município`,
             gender = `Sexo Trabalhador`,
             time_employed = `Tempo Emprego`,
             tp_contract = `Tipo Vínculo`,
             employed_dec = `Vínculo Ativo 31/12`,
             age = `Idade`) %>%
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             date_adm = as.Date(paste0("01/", month_adm, "/", year_adm), format = "%d/%m/%Y"),
             date_quit = as.Date(paste0("30/", month_quit, "/", years_files), format = "%d/%m/%Y"),
             CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root)))) %>% 
      select(-month_adm, -year_adm, -month_quit)
    
  } else if (years_files %in% as.character(c(1999:2001))){
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("CBO 94 Ocupação", "PIS", 
                           "Ano Admissão", "CNPJ Raiz", "Grau Instrução 2005-1985", "Mês Admissão",
                           "Mês Desligamento", "Motivo Desligamento", "Município", "Sexo Trabalhador", 
                           "Tempo Emprego", "Tipo Vínculo", "Vínculo Ativo 31/12", "Idade", "Tipo Admissão")) %>%
      rename(cbo = `CBO 94 Ocupação`,
             year_adm = `Ano Admissão`,
             CNPJ_root = `CNPJ Raiz`,
             educ = `Grau Instrução 2005-1985`,
             month_adm = `Mês Admissão`,
             month_quit = `Mês Desligamento`,
             reason_quit = `Motivo Desligamento`,
             municipality = `Município`,
             gender = `Sexo Trabalhador`,
             time_employed = `Tempo Emprego`,
             tp_contract = `Tipo Vínculo`,
             employed_dec = `Vínculo Ativo 31/12`,
             age = `Idade`)%>%
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             date_adm = as.Date(paste0("01/", month_adm, "/", year_adm), format = "%d/%m/%Y"),
             date_quit = as.Date(paste0("30/", month_quit, "/", years_files), format = "%d/%m/%Y"),
             CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root)))) %>%
      select(-month_adm, -year_adm, -month_quit)
    
  } else{
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("OCUPACAO 94", "PIS", 
                           "HORAS CONTR", "REM DEZ (R$)", "REM MED (R$)", "ULT REM", "SAL CONTR"
                           ,"DT ADMISSAO", "RADIC CNPJ", "GRAU INSTR", "MES DESLIG", "CAUSA DESLI",
                           "MUNICIPIO", "SEXO", "TEMP EMPR", "TP VINCULO", "EMP EM 31/12", "DT NASCIMENT")) %>% 
      rename(cbo = `OCUPACAO 94`,
             date_adm = `DT ADMISSAO`,
             CNPJ_root = `RADIC CNPJ`,
             educ = `GRAU INSTR`,
             month_quit = `MES DESLIG`,
             reason_quit = `CAUSA DESLI`,
             municipality = MUNICIPIO,
             gender = SEXO,
             time_employed = `TEMP EMPR`,
             tp_contract = `TP VINCULO`,
             employed_dec = `EMP EM 31/12`,
             b_date = `DT NASCIMENT`) %>%
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             age = as.numeric(years_files) - as.numeric(substr(b_date,(nchar(b_date)+1)-4,nchar(b_date))),
             date_adm = ifelse(nchar(date_adm) == 7, paste0("0", date_adm), date_adm),
             date_adm = as.Date(date_adm, format = "%d%m%Y"),
             date_quit = as.Date(paste0("30/", month_quit, "/", years_files), format = "%d/%m/%Y"),
             CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root)))) %>%
      select(-b_date, -month_quit)
  }
  
  return(df)
}

####################################################################
#### OPEN FINAL MATCH DF TO GET PIS AND CBO OF EACH OBSERVATION ####
####################################################################
list_PIS_cbo <- list()

for (i in seq_along(years)) {
  list_PIS_cbo[[i]] <- readRDS(paste0(personal_directory, "list_final_df/final_df_", years[i], ".RDS")) %>% 
    select(cbo, PIS) %>% 
    mutate(PIS_cbo = paste0(PIS, "-", cbo)) %>% 
    pull(PIS_cbo)
}

##########################################################
#### OPEN RAIS AND FILTER FOR THE PIS-CBO OF INTEREST ####
##########################################################
list_cnpj <- list()
for(i in seq_along(years)){
  
  print(years[i])
  
  # list files for each year
  files <- grep(list.files(path = paste0(directory, years[i]), 
                           all.files = FALSE,
                           full.names = TRUE), 
                pattern= ('Estb|ESTB|IGNORADO'), 
                inv=TRUE, value=TRUE)
  
  PIS_cbo_vec <- list_PIS_cbo[[i]]
  
  list_data_temp <- list()
  j <- 1
  for(j in seq_along(files)) { 
    # open RAIS file
    list_data_temp[[j]] <- open_files(years_files = years[i], file_to_read = files[j]) %>%
      # filter for corresponding PIS-cbo
      mutate(PIS_cbo = paste0(PIS, "-", cbo)) %>% 
      filter(PIS_cbo %in% PIS_cbo_vec) 
    
    print(j)
  }
  
  #combine all RAIS data
  df_RAIS <- bind_rows(list_data_temp) %>% select(-PIS_cbo)
  
  # open matched data
  df_match <- readRDS(paste0(personal_directory, "list_final_df/final_df_", years[i], ".RDS")) %>% 
    # merge RAIS columns into matched df
    left_join(df_RAIS, by = c("PIS", "cbo"))
  
  # save a "full df" for each year
  saveRDS(df_match, paste0(personal_directory, "list_final_df_full/final_df_full_", years[i], ".RDS"))
  
  # save all CNPJ for each year
  list_cnpj[[i]] <- df_RAIS %>% distinct(CNPJ_root) %>% pull(CNPJ_root)
  
  # clean environment
  remove(df_RAIS, df_match, list_data_temp)
}

#######################################################
#### GETTING CNPJ NAMES FROM RAIS ESTABELECIMENTOS ####
#######################################################
nat_jur_fed <- as.character(c(1015, 1040, 1074, 1104, 1139, 1163, 1252, 10))

years_names_cnpj <- c("2002", "2003", "2004")
# SÓ TEM NOME EM 2002, 2003 e 2004, e parece que só em 2002 está correto
list_cnpj_names <- list()
for(i in seq_along(years_names_cnpj)){
  
  print(years_names_cnpj[i])
  
  # list files for each year
  files <- grep(list.files(path = paste0(directory, years_names_cnpj[i]), 
                           all.files = FALSE,
                           full.names = TRUE), 
                pattern= ('Estb|ESTB'), 
                inv=FALSE, value=TRUE)
  
  list_files <- list()
  for (j in seq_along(files)) {
    list_files[[j]] <- read_delim(files[j], 
                                  "|",
                                  escape_double = FALSE,
                                  col_types = cols(`CLAS CNAE 95` = col_character(), 
                                                   `NATUR JUR` = col_character(), 
                                                   `RADIC CNPJ` = col_character(), 
                                                   `RAZAO SOCIAL` = col_character()),
                                  locale = locale(decimal_mark = ",", grouping_mark = ".", 
                                                  encoding = "WINDOWS-1252"),
                                  trim_ws = TRUE) %>% 
      rename(CNPJ_root = `RADIC CNPJ`,
             agency_name = `RAZAO SOCIAL`,
             cnae95 = `CLAS CNAE 95`,
             natjur = `NATUR JUR`) %>% 
      select(CNPJ_root, agency_name, cnae95, natjur) %>% 
      filter(natjur %in% nat_jur_fed & cnae95 == "75116")
      
  }
  
  list_cnpj_names[[i]] <- bind_rows(list_files) %>% 
    mutate(year = years_names_cnpj[i],
           CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root)))) %>% 
    filter(agency_name != "TESTE") 
  
}

cnpj_names <- bind_rows(list_cnpj_names) %>% 
  count(CNPJ_root_std, agency_name) %>% 
  group_by(CNPJ_root_std) %>% 
  slice(which.max(n)) %>% 
  ungroup()

# adding an old dataset with CNPJ names (from script "analise_paula.Rmd")
rais_panel_public <- readRDS(paste0(directory,"_misc/rais_panel_public.Rds")) %>% 
  distinct(cnpj_raiz_std, name) %>% 
  rename(CNPJ_root_std = cnpj_raiz_std,
         agency_name = name)

CNPJs_filter <- rais_panel_public$CNPJ_root_std

cnpj_names <- cnpj_names %>% 
  select(-n) %>% 
  filter(CNPJ_root_std %notin% CNPJs_filter) %>% 
  bind_rows(rais_panel_public) %>% 
  distinct(CNPJ_root_std, .keep_all = TRUE) %>% 
  mutate(agency_name = ifelse(agency_name == "ISC", "TRIBUNAL DE CONTAS DA UNIAO", agency_name))

write.csv2(cnpj_names, file = paste0(directory, "matching/dec2021/all_cnpj_names.csv"))

################################################
#### INCLUDING CNPJ NAMES IN THE FINAL DATA ####
################################################
for (i in seq_along(years)) {
  df <- readRDS(file = paste0(personal_directory, "list_final_df_full/final_df_full_", years[i], ".RDS")) %>% 
    mutate(CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root)))) %>%
    filter(PIS != "00000000000") %>% 
    distinct(PIS, cbo, CNPJ_root_std, .keep_all = TRUE) %>% 
    left_join(cnpj_names, by = "CNPJ_root_std")
    
  df_cnpj_na <- df %>% filter(is.na(agency_name))
  
  print(nrow(df_cnpj_na %>% distinct(CNPJ_root_std)))
  print(nrow(df %>% distinct(CNPJ_root_std)))
  
  # save a "full df" for each year
  saveRDS(df, paste0(personal_directory, "list_final_df_full/final_df_full_", years[i], ".RDS"))
  saveRDS(df_cnpj_na, paste0(personal_directory, "list_final_df_full/final_df_full_cnpj_na_", years[i], ".RDS"))
}

###############################################################
#### FILTERING ONLY FOR AGENCIES USED IN THE EXPERT SURVEY ####
###############################################################
load(paste0(directory, "matching/dec2021/agencies_manifesto.RData")) 
write.csv2(ag_scores, file = paste0(directory, "matching/dec2021/all_agencies_exp_survey.csv"))

CNPJ_exp_sur <- read.csv2(paste0(directory, "matching/dec2021/all_cnpj_names_expert_survey.csv")) %>% 
  mutate(CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root_std)))) %>% 
  filter(expert_survey == 1) %>% 
  select(-X, -agency_name)

# nrow(CNPJ_exp_sur %>% distinct(agency_name)) # hoje tenho uma base com 70 nomes

agencies <- list()
for (i in seq_along(years)) {
  print(years[i])
  
  df <- readRDS(file = paste0(personal_directory, "list_final_df_full/final_df_full_", years[i], ".RDS")) 
  
  print(nrow(df))
  
  df <- df %>% 
    left_join(CNPJ_exp_sur, by = "CNPJ_root_std") %>% 
    filter(expert_survey == 1)
  
  print(nrow(df))
  print(nrow(df %>% distinct(agency_name)))
  
  CNPJ_df <- df %>% distinct(CNPJ_root_std) %>% pull(CNPJ_root_std)
  
  agencies[[i]] <- CNPJ_exp_sur %>% filter(CNPJ_root_std %in% CNPJ_df) %>% 
    mutate(year = years[i])
  
  # save a "full df" for each year
  saveRDS(df, paste0(personal_directory, "list_final_df_full_exp_survey/final_df_full_exp_sur_", years[i], ".RDS"))
}

CNPJ_exp_sur <- read.csv2(paste0(directory, "matching/dec2021/all_cnpj_names_expert_survey.csv")) %>% 
  mutate(CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root_std)))) %>% 
  filter(expert_survey == 1) %>% 
  select(-X, -expert_survey)

for(i in seq_along(agencies)){
  df <- agencies[[i]] %>% 
    select(-expert_survey)
  
  names(df) <- c("CNPJ_root_std", paste0("year_", years[i]))
  
  CNPJ_exp_sur <- CNPJ_exp_sur %>% 
    left_join(df, by = "CNPJ_root_std")
}

write.csv2(CNPJ_exp_sur, file = paste0(directory, "matching/dec2021/all_cnpj_names_expert_survey_years_v2.csv"))

####################################################
#### EXPORTING ANONIMIZED DATA TO SEND TO PAULA ####
####################################################
pis_vec <- c()

for (i in seq_along(years)) {
  df <- readRDS(paste0(personal_directory, "list_final_df_full_exp_survey/final_df_full_exp_sur_", years[i], ".RDS"))
  
  pis_vec <- unique(c(pis_vec, (df %>% pull(PIS))))
}

set.seed(05585040)

pis_ID <- tibble("PIS" = pis_vec) %>% 
  distinct(PIS) %>% 
  mutate(ID_fake = sample(length(pis_vec)))

for (i in seq_along(years)) {
  df <- readRDS(paste0(personal_directory, "list_final_df_full_exp_survey/final_df_full_exp_sur_", years[i], ".RDS")) %>% 
    left_join(pis_ID, by = "PIS") %>% 
    select(-PIS)
  
  saveRDS(df, paste0(personal_directory, "list_final_df_full_exp_survey_anon/final_df_full_exp_sur_anon_", 
                     years[i], ".RDS"))
}

```

## Commenting script step by step

### Function to open RAIS

After setting some initial definitions, such as directory and necessary packages, I created a function to open the RAIS files.

We are working with the period going from from 1998 to 2004, and unfortunately columns names are not standardized among those years. I defined the necessary columns, look for their names into all df and add it to the function to have standardized df for the rest of the script

The columns are the following:

-   "OCUPACAO": occupation code
-   "PIS": individual identifier
-   "DT ADMISSAO": date when the person started working in this position
-   "RADIC CNPJ": company identifier
-   "GRAU INSTR": educational level of the individual
-   "MES DESLIG": if the person quit/was fired in that year, the month/data when that happened
-   "CAUSA DESLI": the reason why the person quitted/was fired
-   "MUNICIPIO": municipality were the company is located
-   "SEXO": gender
-   "TEMP EMPR": has long the person has been in this position
-   "TP VINCL": type of contract
-   "EMP EM 31/12": if the person was employed (was working in this position) on 31st of December
-   "DT NASCIMENT": birth-date.

```{r open_files, eval = FALSE}
##################################
#### FUNCTION TO OPEN RAIS DF ####
##################################
open_files <- function(years_files = NA_character_, file_to_read = NA_character_){
  if(years_files == "2002"){
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("OCUPACAO", "PIS", 
                           "DT ADMISSAO", "RADIC CNPJ", "GRAU INSTR", "MES DESLIG", "CAUSA DESLI", "MUNICIPIO",
                           "SEXO", "TEMP EMPR", "TP VINCL", "EMP EM 31/12", "DT NASCIMENT")) %>% 
      rename(cbo = `OCUPACAO`,
             date_adm = `DT ADMISSAO`,
             CNPJ_root = `RADIC CNPJ`,
             educ = `GRAU INSTR`,
             month_quit = `MES DESLIG`,
             reason_quit = `CAUSA DESLI`,
             municipality = MUNICIPIO,
             gender = SEXO,
             time_employed = `TEMP EMPR`,
             tp_contract = `TP VINCL`,
             employed_dec = `EMP EM 31/12`,
             b_date = `DT NASCIMENT`) %>%
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             age = as.numeric(years_files) - as.numeric(substr(b_date,(nchar(b_date)+1)-4,nchar(b_date))),
             date_adm = ifelse(nchar(date_adm) == 7, paste0("0", date_adm), date_adm),
             date_adm = as.Date(date_adm, format = "%d%m%Y"),
             date_quit = as.Date(paste0("30/", month_quit, "/", years_files), format = "%d/%m/%Y"),
             CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root)))) %>%
      select(-b_date, -month_quit)
    
  } else if (years_files %in% as.character(1995:1998)){
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("CBO 94 Ocupação", "PIS", 
                           "Ano Admissão", "CNPJ Raiz", "Grau Instrução 2005-1985", "Mês Admissão",
                           "Mês Desligamento", "Motivo Desligamento", "Município", "Sexo Trabalhador", "Tempo Emprego",
                           "Tipo Vínculo", "Vínculo Ativo 31/12", "Idade", "Tipo Admissão")) %>%
      rename(cbo = `CBO 94 Ocupação`,
             year_adm = `Ano Admissão`,
             CNPJ_root = `CNPJ Raiz`,
             educ = `Grau Instrução 2005-1985`,
             month_adm = `Mês Admissão`,
             month_quit = `Mês Desligamento`,
             reason_quit = `Motivo Desligamento`,
             municipality = `Município`,
             gender = `Sexo Trabalhador`,
             time_employed = `Tempo Emprego`,
             tp_contract = `Tipo Vínculo`,
             employed_dec = `Vínculo Ativo 31/12`,
             age = `Idade`) %>%
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             date_adm = as.Date(paste0("01/", month_adm, "/", year_adm), format = "%d/%m/%Y"),
             date_quit = as.Date(paste0("30/", month_quit, "/", years_files), format = "%d/%m/%Y"),
             CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root)))) %>% 
      select(-month_adm, -year_adm, -month_quit)
    
  } else if (years_files %in% as.character(c(1999:2001))){
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("CBO 94 Ocupação", "PIS", 
                           "Ano Admissão", "CNPJ Raiz", "Grau Instrução 2005-1985", "Mês Admissão",
                           "Mês Desligamento", "Motivo Desligamento", "Município", "Sexo Trabalhador", 
                           "Tempo Emprego", "Tipo Vínculo", "Vínculo Ativo 31/12", "Idade", "Tipo Admissão")) %>%
      rename(cbo = `CBO 94 Ocupação`,
             year_adm = `Ano Admissão`,
             CNPJ_root = `CNPJ Raiz`,
             educ = `Grau Instrução 2005-1985`,
             month_adm = `Mês Admissão`,
             month_quit = `Mês Desligamento`,
             reason_quit = `Motivo Desligamento`,
             municipality = `Município`,
             gender = `Sexo Trabalhador`,
             time_employed = `Tempo Emprego`,
             tp_contract = `Tipo Vínculo`,
             employed_dec = `Vínculo Ativo 31/12`,
             age = `Idade`)%>%
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             date_adm = as.Date(paste0("01/", month_adm, "/", year_adm), format = "%d/%m/%Y"),
             date_quit = as.Date(paste0("30/", month_quit, "/", years_files), format = "%d/%m/%Y"),
             CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root)))) %>%
      select(-month_adm, -year_adm, -month_quit)
    
  } else{
    df <- fread(file_to_read, 
                sep=";", 
                header=TRUE, 
                na.strings="NA",
                showProgress = TRUE, 
                colClasses = "character",
                select = c("OCUPACAO 94", "PIS", 
                           "HORAS CONTR", "REM DEZ (R$)", "REM MED (R$)", "ULT REM", "SAL CONTR"
                           ,"DT ADMISSAO", "RADIC CNPJ", "GRAU INSTR", "MES DESLIG", "CAUSA DESLI",
                           "MUNICIPIO", "SEXO", "TEMP EMPR", "TP VINCULO", "EMP EM 31/12", "DT NASCIMENT")) %>% 
      rename(cbo = `OCUPACAO 94`,
             date_adm = `DT ADMISSAO`,
             CNPJ_root = `RADIC CNPJ`,
             educ = `GRAU INSTR`,
             month_quit = `MES DESLIG`,
             reason_quit = `CAUSA DESLI`,
             municipality = MUNICIPIO,
             gender = SEXO,
             time_employed = `TEMP EMPR`,
             tp_contract = `TP VINCULO`,
             employed_dec = `EMP EM 31/12`,
             b_date = `DT NASCIMENT`) %>%
      mutate(cbo = str_replace(cbo, "CBO ", ""),
             age = as.numeric(years_files) - as.numeric(substr(b_date,(nchar(b_date)+1)-4,nchar(b_date))),
             date_adm = ifelse(nchar(date_adm) == 7, paste0("0", date_adm), date_adm),
             date_adm = as.Date(date_adm, format = "%d%m%Y"),
             date_quit = as.Date(paste0("30/", month_quit, "/", years_files), format = "%d/%m/%Y"),
             CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root)))) %>%
      select(-b_date, -month_quit)
  }
  
  return(df)
}
```

### Open data from previous script

```{r}
####################################################################
#### OPEN FINAL MATCH DF TO GET PIS AND CBO OF EACH OBSERVATION ####
####################################################################
list_PIS_cbo <- list()

for (i in seq_along(years)) {
  list_PIS_cbo[[i]] <- readRDS(paste0(personal_directory, "list_final_df/final_df_", years[i], ".RDS")) %>% 
    select(cbo, PIS) %>% 
    mutate(PIS_cbo = paste0(PIS, "-", cbo)) %>% 
    pull(PIS_cbo)
}
```

### Filtering the full df
From the original RAIS data, filter only for the observation included in the matching.
Many observations are not included because they are people that work in the private sector, or they work in the public sector, but in the state or local governmet.

```{r}
##########################################################
#### OPEN RAIS AND FILTER FOR THE PIS-CBO OF INTEREST ####
##########################################################
list_cnpj <- list()
for(i in seq_along(years)){
  
  print(years[i])
  
  # list files for each year
  files <- grep(list.files(path = paste0(directory, years[i]), 
                           all.files = FALSE,
                           full.names = TRUE), 
                pattern= ('Estb|ESTB|IGNORADO'), 
                inv=TRUE, value=TRUE)
  
  PIS_cbo_vec <- list_PIS_cbo[[i]]
  
  list_data_temp <- list()
  j <- 1
  for(j in seq_along(files)) { 
    # open RAIS file
    list_data_temp[[j]] <- open_files(years_files = years[i], file_to_read = files[j]) %>%
      # filter for corresponding PIS-cbo
      mutate(PIS_cbo = paste0(PIS, "-", cbo)) %>% 
      filter(PIS_cbo %in% PIS_cbo_vec) 
    
    print(j)
  }
  
  #combine all RAIS data
  df_RAIS <- bind_rows(list_data_temp) %>% select(-PIS_cbo)
  
  # open matched data
  df_match <- readRDS(paste0(personal_directory, "list_final_df/final_df_", years[i], ".RDS")) %>% 
    # merge RAIS columns into matched df
    left_join(df_RAIS, by = c("PIS", "cbo"))
  
  # save a "full df" for each year
  saveRDS(df_match, paste0(personal_directory, "list_final_df_full/final_df_full_", years[i], ".RDS"))
  
  # save all CNPJ for each year
  list_cnpj[[i]] <- df_RAIS %>% distinct(CNPJ_root) %>% pull(CNPJ_root)
  
  # clean environment
  remove(df_RAIS, df_match, list_data_temp)
}
```

### Including companies names
To make the analysis at agency level, we need to have the company/firm/agency name.
RAIS is divided in two data sets: employees and employers. So far we were working only with the employees data. Now we have to merge it with the employer data set, using the *CNPJ* code. 

First I get the names from the employers data set.
```{r}
#######################################################
#### GETTING CNPJ NAMES FROM RAIS ESTABELECIMENTOS ####
#######################################################
nat_jur_fed <- as.character(c(1015, 1040, 1074, 1104, 1139, 1163, 1252, 10))

years_names_cnpj <- c("2002", "2003", "2004")
# SÓ TEM NOME EM 2002, 2003 e 2004, e parece que só em 2002 está correto
list_cnpj_names <- list()
for(i in seq_along(years_names_cnpj)){
  
  print(years_names_cnpj[i])
  
  # list files for each year
  files <- grep(list.files(path = paste0(directory, years_names_cnpj[i]), 
                           all.files = FALSE,
                           full.names = TRUE), 
                pattern= ('Estb|ESTB'), 
                inv=FALSE, value=TRUE)
  
  list_files <- list()
  for (j in seq_along(files)) {
    list_files[[j]] <- read_delim(files[j], 
                                  "|",
                                  escape_double = FALSE,
                                  col_types = cols(`CLAS CNAE 95` = col_character(), 
                                                   `NATUR JUR` = col_character(), 
                                                   `RADIC CNPJ` = col_character(), 
                                                   `RAZAO SOCIAL` = col_character()),
                                  locale = locale(decimal_mark = ",", grouping_mark = ".", 
                                                  encoding = "WINDOWS-1252"),
                                  trim_ws = TRUE) %>% 
      rename(CNPJ_root = `RADIC CNPJ`,
             agency_name = `RAZAO SOCIAL`,
             cnae95 = `CLAS CNAE 95`,
             natjur = `NATUR JUR`) %>% 
      select(CNPJ_root, agency_name, cnae95, natjur) %>% 
      filter(natjur %in% nat_jur_fed & cnae95 == "75116")
      
  }
  
  list_cnpj_names[[i]] <- bind_rows(list_files) %>% 
    mutate(year = years_names_cnpj[i],
           CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root)))) %>% 
    filter(agency_name != "TESTE") 
  
}

cnpj_names <- bind_rows(list_cnpj_names) %>% 
  count(CNPJ_root_std, agency_name) %>% 
  group_by(CNPJ_root_std) %>% 
  slice(which.max(n)) %>% 
  ungroup()

# adding an old dataset with CNPJ names (from script "analise_paula.Rmd")
rais_panel_public <- readRDS(paste0(directory,"_misc/rais_panel_public.Rds")) %>% 
  distinct(cnpj_raiz_std, name) %>% 
  rename(CNPJ_root_std = cnpj_raiz_std,
         agency_name = name)

CNPJs_filter <- rais_panel_public$CNPJ_root_std

cnpj_names <- cnpj_names %>% 
  select(-n) %>% 
  filter(CNPJ_root_std %notin% CNPJs_filter) %>% 
  bind_rows(rais_panel_public) %>% 
  distinct(CNPJ_root_std, .keep_all = TRUE) %>% 
  mutate(agency_name = ifelse(agency_name == "ISC", "TRIBUNAL DE CONTAS DA UNIAO", agency_name))

write.csv2(cnpj_names, file = paste0(directory, "matching/dec2021/all_cnpj_names.csv"))
```

Then I merge it to the macthcing data set.
```{r}
################################################
#### INCLUDING CNPJ NAMES IN THE FINAL DATA ####
################################################
for (i in seq_along(years)) {
  df <- readRDS(file = paste0(personal_directory, "list_final_df_full/final_df_full_", years[i], ".RDS")) %>% 
    mutate(CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root)))) %>%
    filter(PIS != "00000000000") %>% 
    distinct(PIS, cbo, CNPJ_root_std, .keep_all = TRUE) %>% 
    left_join(cnpj_names, by = "CNPJ_root_std")
    
  df_cnpj_na <- df %>% filter(is.na(agency_name))
  
  print(nrow(df_cnpj_na %>% distinct(CNPJ_root_std)))
  print(nrow(df %>% distinct(CNPJ_root_std)))
  
  # save a "full df" for each year
  saveRDS(df, paste0(personal_directory, "list_final_df_full/final_df_full_", years[i], ".RDS"))
  saveRDS(df_cnpj_na, paste0(personal_directory, "list_final_df_full/final_df_full_cnpj_na_", years[i], ".RDS"))
}
```

### Filtering only for agencies included in the expert survey
```{r}
###############################################################
#### FILTERING ONLY FOR AGENCIES USED IN THE EXPERT SURVEY ####
###############################################################
load(paste0(directory, "matching/dec2021/agencies_manifesto.RData")) 
write.csv2(ag_scores, file = paste0(directory, "matching/dec2021/all_agencies_exp_survey.csv"))

CNPJ_exp_sur <- read.csv2(paste0(directory, "matching/dec2021/all_cnpj_names_expert_survey.csv")) %>% 
  mutate(CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root_std)))) %>% 
  filter(expert_survey == 1) %>% 
  select(-X, -agency_name)

# nrow(CNPJ_exp_sur %>% distinct(agency_name)) # hoje tenho uma base com 70 nomes

agencies <- list()
for (i in seq_along(years)) {
  print(years[i])
  
  df <- readRDS(file = paste0(personal_directory, "list_final_df_full/final_df_full_", years[i], ".RDS")) 
  
  print(nrow(df))
  
  df <- df %>% 
    left_join(CNPJ_exp_sur, by = "CNPJ_root_std") %>% 
    filter(expert_survey == 1)
  
  print(nrow(df))
  print(nrow(df %>% distinct(agency_name)))
  
  CNPJ_df <- df %>% distinct(CNPJ_root_std) %>% pull(CNPJ_root_std)
  
  agencies[[i]] <- CNPJ_exp_sur %>% filter(CNPJ_root_std %in% CNPJ_df) %>% 
    mutate(year = years[i])
  
  # save a "full df" for each year
  saveRDS(df, paste0(personal_directory, "list_final_df_full_exp_survey/final_df_full_exp_sur_", years[i], ".RDS"))
}

CNPJ_exp_sur <- read.csv2(paste0(directory, "matching/dec2021/all_cnpj_names_expert_survey.csv")) %>% 
  mutate(CNPJ_root_std = base::sprintf("%08.0f", as.numeric(as.character(CNPJ_root_std)))) %>% 
  filter(expert_survey == 1) %>% 
  select(-X, -expert_survey)

for(i in seq_along(agencies)){
  df <- agencies[[i]] %>% 
    select(-expert_survey)
  
  names(df) <- c("CNPJ_root_std", paste0("year_", years[i]))
  
  CNPJ_exp_sur <- CNPJ_exp_sur %>% 
    left_join(df, by = "CNPJ_root_std")
}

write.csv2(CNPJ_exp_sur, file = paste0(directory, "matching/dec2021/all_cnpj_names_expert_survey_years_v2.csv"))
```

### Ananimizing and exporting
```{r}
####################################################
#### EXPORTING ANONIMIZED DATA TO SEND TO PAULA ####
####################################################
pis_vec <- c()

for (i in seq_along(years)) {
  df <- readRDS(paste0(personal_directory, "list_final_df_full_exp_survey/final_df_full_exp_sur_", years[i], ".RDS"))
  
  pis_vec <- unique(c(pis_vec, (df %>% pull(PIS))))
}

set.seed(05585040)

pis_ID <- tibble("PIS" = pis_vec) %>% 
  distinct(PIS) %>% 
  mutate(ID_fake = sample(length(pis_vec)))

for (i in seq_along(years)) {
  df <- readRDS(paste0(personal_directory, "list_final_df_full_exp_survey/final_df_full_exp_sur_", years[i], ".RDS")) %>% 
    left_join(pis_ID, by = "PIS") %>% 
    select(-PIS)
  
  saveRDS(df, paste0(personal_directory, "list_final_df_full_exp_survey_anon/final_df_full_exp_sur_anon_", 
                     years[i], ".RDS"))
}
```

